### 2016-11-26 discussion.tex
added a start to some sort of writeup and a look at the simulation.... Not sure where that is going

\documentclass[11pt]{article}
\usepackage{fullpage,amsmath,amsfonts,mathpazo,microtype,nicefrac,graphicx,verbatimbox,listings,hyperref,enumitem,amssymb,float,pdfpages}
% Macro definitions
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\B}{\mathbb{B}}
\newcommand{\p}{\partial}
\newcommand{\Trans}{\mathsf{T}}
\renewcommand{\vec}[1]{\mathbf{#1}}
\newcommand{\vx}{\vec{x}}
\newcommand{\vb}{\vec{b}}

\DeclareMathOperator{\rank}{rank}

% Include support for including a pdf

% \begin{lstlisting}[frame=single]
% \end{lstlisting}

% \begin{figure}[H]
% \centering
% \includegraphics[keepaspectratio=true,scale=0.6]{problem3d}
% \caption{Plot of temperature cross section in pipe at various time-steps of simulation}\label{visina8}
% \end{figure}

\begin{document}

\lstset{language=Python, basicstyle=\scriptsize} % set language as python

\section*{AM205 Project - Traffic Flow Modelling; do autonomous cars help? (I don't know about this but we can change it)}

\section{Introduction}
The modeling of traffic at traffic lights, on highways, in dense streams etc. can often present a problem that is too complicated to solve analytically. This is mainly due to the inherent random movements of cars, the number of individually moving cars and the many scenarios that can be analyzed. These scenarios can be modeled numerically by applying the equations of motion to a system of vehicles and observing the impact of a number of outcomes. In this project, we will present an analytical solution to a number of simplistic scenarios involving traffic (cars starting from a red light, cars approaching a red light etc). We will then present the same scenario using a numerically calculated solution, modeling the cars using a differential setting where one car's motion is dependent on the car that is in front of it. We will then extend the scope of this analysis and introduce a number of varying factors into the model to analyze the motion of cars along a straight road, analyze the cars along a circular track and finally include autonomous cars with the human controlled vehicles. These autonomous cars have more precise, and faster reacting velocity controllers and the inclusion of vehicles such as these in the model is expected to smooth the flow of traffic and ultimately increase the density of traffic flow.

\section{Microscopic Perspective of Traffic Interaction - The Analytical Solution}

\section{Macroscopic Perspective of Traffic Interaction}
For this section of the problem, we will model the behaviors of cars in a stream of traffic individually. We will then simulate the movement of this stream of cars and demonstrate some well known traffic phenomenon (cars starting from a red traffic light will propagate forward in a wave formation, cars traveling along a straight road will oscillate into a traffic jam). Furthermore, we will use different controllers to simulate the means by which cars can control their acceleration and will be able to simulate the effect that autonomous cars may have for the throughput of a stream of traffic for varying car densities. This numerical analysis can be contrasted to the analytical solution that was presented in (**another section 2?**).\par
A cars speed is governed by the laws of motion:
\begin{equation}
v_f = v_i + a\Delta t
\end{equation}

An individual car can be modeled as such over a 100,000m track and asked to travel at $v_{target} = 60miles.h^{-1}$ (i.e. $26.8m.s^{-1}$). As, the driver is not an perfect controller he is unable to hold the speed exactly at $v_{target}$. We can therefore model his fluctuations as random acceleration with the corresponding adjustments to correct the acceleration:
\begin{equation}
a_{correction_i} = \alpha(v_{instant} - v_{target}) + N(0,\sigma_1)
\end{equation}

We note that for more than one car in a stream of traffic the car's acceleration is then also dependent on the car ahead of it:
\begin{equation}\label{eq:interaction}
a_{interaction_i} = \beta (s_{instant} - s_{target})
\end{equation}

In equation \ref{eq:interaction}, we see that the acceleration of the car is also dependent on the distance between any one car and the car that is directly ahead of it. The final acceleration of the car is therefore:
\begin{equation}\label{eq:acceleration}
a_{i} = a_{correction_i} + a_{interaction_i}
\end{equation}

If we assume the cars should travel at $v_{target}$ and that they should maintain an $s_{target} = 5m$ spacing between them, we can then model the velocity of the car as:

\begin{equation}\label{eq:velocity}
v_{i} = \frac{d(a_i)}{dt}
\end{equation}

\end{document}

### 2016-11-26 code/simulation.py

@@ -1,0 +1,204 @@
+"""
+A simple example of an animated plot
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.animation as animation
+from scipy.integrate import odeint
+import pdb
+
+
+n_cars = 5 # use later
+meters = 100000
+v_target = 26.8 # 60miles/h in m/s
+d_target = 5 # ideal spacing of 5m between cars
+# velocity = v_target
+
+# timestep = 1.
+# distance = 0
+# gain = -2 # strength of controller
+# acceleration = 0
+# time = 0
+# distances = []
+# velocities = []
+# times = []
+
+# while distance < meters:
+# 	distances.append(distance)
+# 	velocities.append(velocity-v_target)
+# 	times.append(time)
+
+# 	time += timestep
+# 	distance += velocity*timestep + 0.5*acceleration*timestep**2
+# 	velocity = velocity + acceleration*timestep
+# 	# try and control for the acceleration by the velocity readings
+# 	acceleration = gain*(velocity - v_target)/timestep
+# 	# add some noise to acceleration
+# 	acceleration = acceleration + np.random.normal(0,0.1)
+#!/usr/bin/python
+
+#########################################
+# Rycroft model
+
+# Constants in model
+# alpha=1.5
+# beta=0.5
+# gamma=0.4
+# delta=0.4
+
+# Function
+# def deriv(x,t):
+#   return np.array([alpha*x[0]-beta*x[0]*x[1],-gamma*x[1]+delta*x[0]*x[1]])
+
+# Solve ODE using the "odeint" library in SciPy
+# time=np.linspace(0,70,500)
+
+# # Initial conditions, set to the initial
+# xinit=np.array([10,5])
+# x=odeint(deriv,xinit,time)
+
+# for i in range(0,500):
+#     print time[i],x[i,0],x[i,1]
+
+# # Plot the solutions
+# plt.figure()
+# p0,=plt.plot(time,x[:,0])
+# p1,=plt.plot(time,x[:,1])
+# plt.legend([p0,p1],["B(t)","C(t)"])
+# plt.xlabel('t')
+# plt.ylabel('Population')
+# plt.show()
+########################################
+
+velocity = np.ones(n_cars) * v_target
+distance = np.arange(n_cars)*5 # everyone starts 5m apart
+
[…]

### 2016-12-05 code/human_driver_model.py
Fixed HDM bug

@@ -100,13 +100,9 @@ def x_v_dash2(x_v, t,params,past):
 	# put s for car zero within the bounds of the track
 	s[0] += end_of_track
 	# Compute index of this timestep
-	index = math.floor(t/t_step) if math.floor(t/t_step) < t_steps else t_steps-1
+	index = int(math.floor(t/t_step)) if math.floor(t/t_step) < t_steps else t_steps-1
 	# compute estimate of gap
 	s_est = s * np.exp(Vs * w_s[:, index])
-	# Update history
-	#params['past_s_s'].append(s)
-	#params['past_s_est_s'].append(s_est)
-
 	# Compute estimates v^est
 	# Note that for vehicle i, v^est_l[i] is vehicle i's estimate of
 	# vehicle i-1's speed
@@ -120,7 +116,6 @@ def x_v_dash2(x_v, t,params,past):
 	delta_v_est = v - v_l_est
 	past['past_delta_v_est_s'].append(delta_v_est)
 	# update history
-	#past_v_s = np.array(params['past_v_s'])
 	past_v_l_est_s = np.array(past['past_v_l_est_s'])
 	past_v_est_s = np.roll(past_v_l_est_s,-1)
 	past_dvdt = np.array(past['past_dvdt'])
@@ -133,9 +128,10 @@ def x_v_dash2(x_v, t,params,past):
 		v_l_prog = interp(past_v_l_est_s,0,params)
 		v_prog = interp(past_v_est_s,0,params) + Tr * interp(past_dvdt,0,params)
 		# compute s^prog
+		#print s_est
 		s_prog = s_est - Tr * interp(past_delta_v_est_s,0,params)
 		# compute c_idm
-		c_idm = np.sum([1./j**2 for j in range(1, n_a+1)])**-1
+		c_idm = np.sum([1./j**2 for j in xrange(1, n_a+1)])**(-1)
 		free_term = a_IDM_free(v, params)
 		dvdt = np.zeros(v.shape)
 		# calculate the acceleration of each vehicle one at a time
@@ -143,14 +139,17 @@ def x_v_dash2(x_v, t,params,past):
 		for alpha in xrange(n_cars):
 			free_term_alpha = free_term[alpha]
 			int_term = 0.0
+			v_alpha_prog = np.array([v_prog[alpha]])
 			for beta in xrange(alpha-n_a,alpha):
-				s_alpha_beta_prog = np.array([np.sum(s_prog[beta+1:alpha+1])])
[…]

### 2016-12-05 code/intelligent_driver_model.py

@@ -31,6 +31,8 @@ def a_IDM(v,s,delta_v,params):
 	delta = params['delta']
 	v0 = params['v0']
 	# Compute dv/dt (i.e. accelerations)
+	#print v0
+	#print s
 	return a*(1-(v/v0)**delta - (s_star(v, delta_v, params)/s)**2)
 
 def compute_a_free(v,s,delta_v,params):

### 2016-12-05 code/human_driver_model.py
Added figures for HDM anticipating 1,3,5 cars

@@ -189,7 +189,6 @@ if __name__ == '__main__':
 	params['total_time'] = 500 # total time (in s)
 	params['t_step'] = (params['total_time'] - params['t_start'])/params['t_steps']
 	params['Tr'] = 0.6 # Reaction time
-	params['na'] = 5 	# number of look ahead cars
 	params['Vs'] = 0.1 # Variation coefficient of gap estimation error
 	params['sigma_r'] = 0.01 # estimation error for the inverse TTC
 	params['sigma_a']= 0.1  # magnitude of acceleration noise
@@ -207,7 +206,6 @@ if __name__ == '__main__':
 	n_cars = params['n_cars']
 	total_time = params['total_time']
 	Tr = params['Tr']
-	na = params['na']
 	Vs = params['Vs']
 	sigma_r = params['sigma_r']
 	sigma_a = params['sigma_a']
@@ -215,7 +213,7 @@ if __name__ == '__main__':
 	tau_a_tilde = params['tau_a_tilde']
 	t_start = params['t_start']
 	t_step = params['t_step']
-	params['n_a'] = 5 # number of cars ahead that the driver is aware of
+	params['n_a'] = n_a # number of cars ahead that the driver is aware of
 	params['w_s'] = wiener_process(tau_tilde, params)
 	params['w_l'] = wiener_process(tau_tilde, params)
 	params['w_a'] = wiener_process(tau_a_tilde, params)
@@ -257,11 +255,10 @@ if __name__ == '__main__':
 	axes[0].set_ylabel('Position')
 	axes[1].set_xlabel('Time')
 	axes[1].set_ylabel('Velocity')
-	plot_out_name = 'HDM_anticipate_{}'.format(params['n_a'])
+	plot_out_name = '../figures/HDM_anticipate_{}.pdf'.format(params['n_a'])
 	plt.savefig(plot_out_name,
 				orientation='landscape',format='pdf',edgecolor='black')
-		plt.close()
-
+	plt.close()
 	# Run a simulation of sorts
 	# Plot Animation of cars on ring track
 	r = (end_of_track/(2*np.pi))

### 2016-12-06 code/blended_models_2.py
only kind of working

@@ -1,0 +1,355 @@
+# Human Driver Model
+import numpy as np
+import scipy as sp
+import scipy.integrate
+import matplotlib
+import matplotlib.pyplot as plt
+import scipy.stats
+import math
+# import scipy.stats.norm
+import matplotlib.animation as animation
+import pdb
+from intelligent_driver_model import *
+from human_driver_model import *
+from collections import deque
+
+def compute_a_free(v,s,delta_v,params):
+	# compute free acceleration
+	a = params['a_idm']
+	delta = params['delta']
+	v0 = params['v0_idm']
+	b = params['b']
+	mask = (v<=v0)
+	# pdb.set_trace()
+	return mask*(a*(1-(v/v0)**delta))+(1-mask)*(-b*(1-(v0/v)**(a*delta/b)))
+
+def a_IIDM(v,s,delta_v,params):
+	# compute acceleration for IIDM model
+	a = params['a_idm']
+	v0 = params['v0_idm']
+	# compute z
+	z = s_star(v, delta_v, params)/s
+	v_mask = (v<=v0)
+	z_mask = (z>=1)
+	a_free = compute_a_free(v,s,delta_v,params)
+	dvdt = np.zeros(v.shape)
+	# v <= v0 and z >= 1
+	dvdt += v_mask * z_mask * (a*(1-z**2))
+	# v <= v0 and z < 1
+	dvdt += v_mask * (1-z_mask) * (a_free*(1-z**(2*a/a_free)))
+	# v > v0 and z >= 1
+	dvdt += (1-v_mask) * z_mask * (a_free+a*(1-z**2))
+	# v > v0 and z < 1
+	dvdt += (1-v_mask) * z_mask * (a_free)
+	return dvdt
+
+
+def x_v_dash(x_v, t,indices, params,past):
+	# Compute derivative of position and velocity
+	x_v = x_v.reshape(2,-2)
+	# get velocities
+	v = x_v[1,:]
+	# get positions
+	x_vec = x_v[0,:]
+	## Follow the leader ##
+	# Note: Car i follows car i-1
+	# Since this is a ring track, car 0 follows car n-1
+	# for vehicle i, delta_v = v[i] - v[i-1]
+	vl = np.roll(v,1)
+
+	# Compute difference in speeds between current car and leading car
+	delta_v = v - vl
+
+	next_v_l_est = np.zeros(params['n_cars'])
[…]

### 2016-12-06 code/blended_models_2.py
fixed bug

@@ -23,6 +23,11 @@ def compute_a_free(v,s,delta_v,params):
 	# pdb.set_trace()
 	return mask*(a*(1-(v/v0)**delta))+(1-mask)*(-b*(1-(v0/v)**(a*delta/b)))
 
+def a_IDM_int(v,s,delta_v,params):
+	# Compute a^IDM_int
+	a = params['a_idm']
+	return -a*(s_star(v, delta_v, params)/s)**2
+
 def a_IIDM(v,s,delta_v,params):
 	# compute acceleration for IIDM model
 	a = params['a_idm']
@@ -165,7 +170,7 @@ def x_v_dash2(x_v, t,indices, params,past):
 		dvdt = np.zeros(v.shape)
 		# calculate the acceleration of each vehicle one at a time
 		# As written in Eq 12.20 of the textbook
-		for alpha in xrange(n_HDM_cars):
+		for alpha in xrange(n_cars):
 			free_term_alpha = free_term[alpha]
 			int_term = 0.0
 			v_alpha_prog = np.array([v_prog[alpha]])
@@ -218,7 +223,7 @@ def runge_kutta_blended_4(x_v_vec_k,x_v_dash, x_v_dash_2, t_k, h,indices1, indic
 if __name__ == '__main__':
 	## Parameters ##
 	params = dict()
-	params['v0'] = 250.0 # desired velocity (in m/s) of vehicles in free traffic
+	params['v0'] = 150.0 # desired velocity (in m/s) of vehicles in free traffic
 	params['init_v'] = 5.0 # initial velocity
 	params['T'] = 1. # Safe following time
 	# Maximum acceleration (in m/s^2)
@@ -228,18 +233,18 @@ if __name__ == '__main__':
 	# SEE PAGE 216 in the book
 	# Ex: a=2.0 is unstable
 	# Ex: a=1.0 is stable
-	params['a'] = 1.0
+	params['a'] = 2.0
 	params['T_idm'] = 1.
-	params['a_idm'] = 1.0
+	params['a_idm'] = 2.0
 	params['v0_idm'] = 150.0
 	params['b'] = 3.0 # Comfortable deceleration (in m/s^2)
 	params['delta'] = 4.0 # Acceleration exponent
 	params['s0'] = 2.0 # minimum gap (in m)
 	params['end_of_track'] = 600 # in m
-	params['t_steps'] = 5000 # number of timesteps
+	params['t_steps'] = 1000 # number of timesteps
 	params['t_start'] = 0.0
[…]

### 2016-12-06 code/blended_models_2.py
fucking with parameters to get some plots. Revert this commit if shit is breaking

@@ -82,7 +82,7 @@ def x_v_dash(x_v, t,indices, params,past):
 	# put s for car zero within the bounds of the track
 	s = np.roll(x_vec,1) - x_vec
 	s[0] += params['end_of_track']
-
+	s[0] = 1000
 	# Compute acceleration
 	if params['IDM_model_num'] == 0:
 		# Standard IDM
@@ -127,6 +127,7 @@ def x_v_dash2(x_v, t,indices, params,past):
 	s = np.roll(x_vec,1) - x_vec
 	# put s for car zero within the bounds of the track
 	s[0] += end_of_track
+	s[0] = 1000
 	# Compute index of this timestep
 	index = int(math.floor(t/t_step)) if math.floor(t/t_step) < t_steps else t_steps-1
 	# compute estimate of gap
@@ -146,6 +147,7 @@ def x_v_dash2(x_v, t,indices, params,past):
 	# Since this is a ring track, car 0 follows car n-1
 	# for vehicle i, delta_v_est = v[i] - v_est[i-1]
 	delta_v_est = v - v_l_est
+	# delta_v_est[0] = float('inf')
 	# next_delta_v_est = np.zeros(n_cars*2)
 	# next_delta_v_est[indices] = delta_v_est
 	past['past_delta_v_est_s'][-1][indices[:len(indices)//2]] = delta_v_est[indices[:len(indices)//2]]
@@ -187,6 +189,15 @@ def x_v_dash2(x_v, t,indices, params,past):
 	else:
 		# Compute acceleration (with estimation error)
 		dvdt = a_IDM(v,s_est,delta_v_est,params) + sigma_a*w_a[:, index]
+
+	# if t <= 20:
+	# 	v[0] = 15
+	if t > 25:
+		dvdt[0] = .5 + sigma_a*w_a[0, index]
+	if t > 30:
+		dvdt[0] = -1.5 + sigma_a*w_a[0, index]
+	if t > 45:
+		dvdt[0] =  sigma_a*w_a[0, index]
 	x_v = np.concatenate((v,dvdt))
 	past['past_dvdt'][-1][indices[:len(indices)//2]] = dvdt[indices[:len(indices)//2]]
 	return x_v
@@ -225,7 +236,7 @@ if __name__ == '__main__':
 	params = dict()
 	params['v0'] = 150.0 # desired velocity (in m/s) of vehicles in free traffic
 	params['init_v'] = 5.0 # initial velocity
-	params['T'] = 1. # Safe following time
[…]

### 2016-12-06 code/intelligent_driver_model.py

@@ -96,6 +96,7 @@ def x_v_dash(x_v, t, params):
 	# Since this is a ring track, car 0 follows car n-1
 	# for vehicle i, delta_v = v[i] - v[i-1]
 	vl = np.roll(v,1)
+	vl[0] = float('inf')
 	# Compute difference in speeds between current car and leading car
 	delta_v = v - vl
 
@@ -105,6 +106,7 @@ def x_v_dash(x_v, t, params):
 	s = np.roll(x_vec,1) - x_vec
 	s[0] += params['end_of_track']
 
+	s[0] = float('inf')
 	# Compute acceleration
 	if params['IDM_model_num'] == 0:
 		# Standard IDM
@@ -117,9 +119,16 @@ def x_v_dash(x_v, t, params):
 	if params['IDM_model_num'] == 2:
 		# AAC IDM
 		dvdt = a_ACC(s, v, vl, dvdt, params)
-
+	if t <= 200:
+		v[0] = 15
+	if t > 200:
+		dvdt[0] = 1
+	if t > 240:
+		dvdt[0] = -1
+	if t > 280:
+		dvdt[0] = 0
 	x_v = np.concatenate((v,dvdt))
-	#params['y_s'].append(x_v)
+	# params['y_s'].append(x_v)
 	return x_v
 
 def runge_kutta_5(x_v_vec_k, x_v_dash, t_k, h, params):
@@ -130,6 +139,7 @@ def runge_kutta_5(x_v_vec_k, x_v_dash, t_k, h, params):
 	k5 = x_v_dash(x_v_vec_k + 439./216*h*k1 -8.*h*k2 + 3680./513*h*k3 - 845./4104*h*k4,t_k*h, params)
 	k6 = x_v_dash(x_v_vec_k - 8./27*h*k1 + 2.*h*k2 -3544./2565*h*k3 + 1859./4104*h*k4 - 11./40*h*k5,t_k*1./2*h, params)
 	x_v_vec_k_next = x_v_vec_k + h * (15./135*k1 + 6656./12825*k3 + 28561./56430*k4 - 9./50*k5 + 2./55*k6)
+	params['x_v_dash'].append(k1)
 	return x_v_vec_k_next
 
 def runge_kutta_4(x_v_vec_k, x_v_dash, t_k, h, params):
[…]

### 2016-12-08 discussion.tex
report with tentative section headers and todo list

@@
 
+\section{Macroscopic Perspective of Traffic Interaction}
+There are generally two ways to model traffic.
+In the macroscopic perspective, traffic is viewed as a fluid or gas with a given maximum density moving according to the laws of conservation of mass.
+Because we'd like to interweave different models for specific cars, these models would be difficult to blend in a way that is easily interpretable.
+
 \section{Microscopic Perspective of Traffic Interaction - The Analytical Solution}
 
-\section{Macroscopic Perspective of Traffic Interaction}
+\section{Microscopic Perspective of Traffic Interaction : \\Intelligent Driver Models}
+In the microscopic perspective, individual cars are modeled as particles that move according to a relationship to their leading particle(s).
 For this section of the problem, we will model the behaviors of cars in a stream of traffic individually.
@@
 This numerical analysis can be contrasted to the analytical solution that was presented in (**another section 2?**).\par
-A cars speed is governed by the laws of motion:
+A car`s speed is governed by the laws of motion:
 \begin{equation}
@@
 $26.8m.s^{-1}$).
-As, the driver is not an perfect controller he is unable to hold the speed exactly at $v_{target}$.
+As the driver is not an perfect controller he is unable to hold the speed exactly at $v_{target}$.
 We can therefore model his fluctuations as random acceleration with the corresponding adjustments to correct the acceleration:
@@
 
+\section{Microscopic Perspective of Traffic Interaction : Human Driver Models}
+
+\section{Numerical Simulations of Traffic Systems}
+
+\section{Stability Comparison of a Variety of Numerical Integration Schemes}
+
+\section{String Stability of Traffic Systems to Perturbation}
+
+\section{Optimization of Traffic Systems}
+
+\section{TODO}
+run stability analysis of Runge Kutta schemes \\
+develop a way to run smaller timesteps over any portion we think is likely to explode (although not actually sure how much this will help) \\
+develop a cost function for emissions/energy usage and optimize over that \\
+can we find the parameters that will cause a phantom traffic jam?
+
 \end{document}

### 2016-12-09 code/HDM.py
differencing and trying difft parameters

@@ -1,0 +1,72 @@
+#https://en.wikipedia.org/wiki/Intelligent_driver_model
+
+
+import numpy as np
+import scipy as sp
+import scipy.integrate
+import matplotlib
+import matplotlib.pyplot as plt
+import pdb
+
+v0 = 30.
+T = 1.5
+a = 1.
+b = 3.
+delta = 4.
+s0 = 2.
+end_of_track = 600
+t_steps = 100000
+
+
+def x_dash(t, v):
+	return v
+
+def x(x0, v, t):
+	return x0 + v*t
+
+def s_star(v, delta_v):
+	return s0 + v*T + (v*delta_v/(2*np.sqrt(a*b)))**2
+
+def x_v_dash(x_v, t):
+	# pdb.set_trace()
+	x_v = x_v.reshape(2,-2)
+	v = x_v[1,:]
+	x_vec = x_v[0,:]
+	delta_v = v.copy()
+
+	s = x_vec.copy()
+
+	for i in range(len(s)):
+		if i == len(s)-1:
+			delta_v[i] = v[0] - v[-1]
+			s[i] =  -end_of_track%(x_vec[0] - x_vec[-1])
+		else:
+			delta_v[i] =  v[i] - v[i+1]
+			s[i] =  x_vec[i] - x_vec[i+1]
+	# print x_vec
+	# print s
+	# print delta_v
+	# if (x_vec[-1] < 0.1):
+	# pdb.set_trace()
+	x_v = np.concatenate(([v], [a*(1-(v/v0)**delta - (s_star(v, delta_v)/s)**2)]), axis=0).reshape(1,-1)[0]
+	return x_v
+
+v = np.ones(50) * 30
+x_vec = np.linspace(0,490,50)
+x_v_vec = np.concatenate(([x_vec], [v]), axis=0).reshape(1,-1)[0]
+
+ts = np.linspace(0,1000,t_steps)
+
+y_s = sp.integrate.odeint(x_v_dash, y0=x_v_vec, t=ts)
+
+fig, ax = plt.subplots(1,1, figsize=(8,8))
+# pdb.set_trace()q
+y_s = y_s.reshape(t_steps,-2,2)
+# for car in range(50):
+# 	ax.plot(ts, x_vec)
+for car in range(50):
+	# pdb.set_trace()
+	print y_s[:,car,0]
+	ax.plot(ts, y_s[:,car,0])
+# print xs
+plt.show()

### 2016-12-09 code/blended_models.py

@@ -9,30 +9,44 @@ import math
 # import scipy.stats.norm
 import matplotlib.animation as animation
 import pdb
-from intelligent_driver_model import *
-from human_driver_model import *
+from intelligent_driver_model import s_star, x_dash, x
+# from human_driver_model import *
 from collections import deque
 import warnings
 
-def compute_a_free(v,s,delta_v,params):
-	# compute free acceleration
-	a = params['a_idm']
+def interp(u,i,params):
+	j = params['j']
+	r = params['r']
+	u_interp = r*u[i-j-1] + (1-r)*u[i-j]
+	return u_interp
+
+def wiener_process(tau_tilde, params):
+	dt = params['t_step']
+	n_cars = params['n_cars']
+	t_steps = params['t_steps']
+	w0 = np.random.randn(n_cars)
+	w = np.zeros((n_cars, t_steps))
+	w[:, 0] = w0
+
+	for time_step in range(1, int(t_steps)):
+		w[:, time_step] = np.exp(-dt/tau_tilde)*w[:, time_step-1] + np.sqrt(2*dt/tau_tilde)*np.random.randn(n_cars)
+
+	return w
+
+def a_IDM(v,s,delta_v,params):
+	# compute acceleration for IDM model
+	a = params['a']
 	delta = params['delta']
-	v0 = params['v0_idm']
-	b = params['b']
-	mask = (v<=v0)
-	# pdb.set_trace()
-	return mask*(a*(1-(v/v0)**delta))+(1-mask)*(-b*(1-(v0/v)**(a*delta/b)))
+	v0 = params['v0']
+	# Compute dv/dt (i.e. accelerations)
+	#print v0
+	#print s
+	return a*(1-(v/v0)**delta - (s_star(v, delta_v, params)/s)**2)
 
-def a_IDM_int(v,s,delta_v,params):
-	# Compute a^IDM_int
-	a = params['a_idm']
-	return -a*(s_star(v, delta_v, params)/s)**2
-
-def a_IIDM(v,s,delta_v,params):
+def a_IIDM(v,s,delta_v,indices, params):
 	# compute acceleration for IIDM model
-	a = params['a_idm']
-	v0 = params['v0_idm']
+	a = params['a']
[…]

### 2016-12-09 code/human_driver_model.py

@@ -178,7 +178,7 @@ if __name__ == '__main__':
 	# SEE PAGE 216 in the book
 	# Ex: a=2.0 is unstable
 	# Ex: a=1.0 is stable
-	params['a'] = 1.0 
+	params['a'] = 1.0
 	params['b'] = 3.0 # Comfortable deceleration (in m/s^2)
 	params['delta'] = 4.0 # Acceleration exponent
 	params['s0'] = 2.0 # minimum gap (in m)
@@ -213,7 +213,7 @@ if __name__ == '__main__':
 	tau_a_tilde = params['tau_a_tilde']
 	t_start = params['t_start']
 	t_step = params['t_step']
-	params['n_a'] = n_a # number of cars ahead that the driver is aware of
+	params['n_a'] = 5 # number of cars ahead that the driver is aware of
 	params['w_s'] = wiener_process(tau_tilde, params)
 	params['w_l'] = wiener_process(tau_tilde, params)
 	params['w_a'] = wiener_process(tau_a_tilde, params)
@@ -255,9 +255,9 @@ if __name__ == '__main__':
 	axes[0].set_ylabel('Position')
 	axes[1].set_xlabel('Time')
 	axes[1].set_ylabel('Velocity')
-	plot_out_name = '../figures/HDM_anticipate_{}.pdf'.format(params['n_a'])
+	plot_out_name = '../figures/HDM_anticipate_{}.png'.format(params['n_a'])
 	plt.savefig(plot_out_name,
-				orientation='landscape',format='pdf',edgecolor='black')
+				orientation='landscape',format='png',edgecolor='black')
 	plt.close()
 	# Run a simulation of sorts
 	# Plot Animation of cars on ring track

### 2016-12-09 code/intelligent_driver_model.py

@@ -96,7 +96,7 @@ def x_v_dash(x_v, t, params):
 	# Since this is a ring track, car 0 follows car n-1
 	# for vehicle i, delta_v = v[i] - v[i-1]
 	vl = np.roll(v,1)
-	vl[0] = float('inf')
+	# vl[0] = float('inf')
 	# Compute difference in speeds between current car and leading car
 	delta_v = v - vl
 
@@ -106,7 +106,7 @@ def x_v_dash(x_v, t, params):
 	s = np.roll(x_vec,1) - x_vec
 	s[0] += params['end_of_track']
 
-	s[0] = float('inf')
+	# s[0] = float('inf')
 	# Compute acceleration
 	if params['IDM_model_num'] == 0:
 		# Standard IDM
@@ -119,14 +119,14 @@ def x_v_dash(x_v, t, params):
 	if params['IDM_model_num'] == 2:
 		# AAC IDM
 		dvdt = a_ACC(s, v, vl, dvdt, params)
-	if t <= 200:
-		v[0] = 15
-	if t > 200:
-		dvdt[0] = 1
-	if t > 240:
-		dvdt[0] = -1
-	if t > 280:
-		dvdt[0] = 0
+	# if t <= 200:
+	# 	v[0] = 15
+	# if t > 200:
+	# 	dvdt[0] = 1
+	# if t > 240:
+	# 	dvdt[0] = -1
+	# if t > 280:
+	# 	dvdt[0] = 0
 	x_v = np.concatenate((v,dvdt))
 	# params['y_s'].append(x_v)
 	return x_v
@@ -250,6 +250,7 @@ if __name__ == '__main__':
 		for i in range(1,len(ts)):
 			y_s.append(runge_kutta_3(y_s[-1], x_v_dash, ts[i], ts[i]-ts[i-1], params))
 
+		y_s = np.array(y_s)
 		# Plot position and velocity of each car
 		fig, axes = plt.subplots(1,2, figsize=(16,8))
 		# Plot positions over time
@@ -262,9 +263,9 @@ if __name__ == '__main__':
 		axes[0].set_ylabel('Displacement')
 		axes[1].set_xlabel('Time')
 		axes[1].set_ylabel('Velocity')
-		plot_out_name = "../figures/displacement_and_velocity_plot_model{}.pdf".format(params['IDM_model_num'])
+		plot_out_name = "../figures/displacement_and_velocity_plot_model{}.png".format(params['IDM_model_num'])
 		plt.savefig(plot_out_name,
[…]

### 2016-12-09 code/test_finite_differences.py

@@ -1,0 +1,88 @@
+# Human Driver Model
+import numpy as np
+import scipy as sp
+import scipy.integrate
+import matplotlib
+import matplotlib.pyplot as plt
+import scipy.stats
+import math
+# import scipy.stats.norm
+import matplotlib.animation as animation
+import pdb
+from intelligent_driver_model import *
+# from human_driver_model import *
+from collections import deque
+
+if __name__ == '__main__':
+	## Parameters ##
+    params 									= dict()
+    params['v0'] 						= 30.0 # desired velocity (in m/s) of vehicles in free traffic
+    params['init_v'] 				= 5.0 # initial velocity
+    params['T'] 						= 1.0 # Safe following time
+    params['a'] 						= 2.0 # Maximum acceleration (in m/s^2)
+    params['b'] 						= 3.0 # Comfortable deceleration (in m/s^2)
+    params['delta'] 				= 4.0 # Acceleration exponent
+    params['s0'] 						= 2.0 # minimum gap (in m)
+    params['end_of_track'] 	= 600 # in m
+    params['n_cars'] 				= 50 # number of vehicles
+    params['total_time'] 		= 500 # total time (in s)
+    params['c'] 						= 0.99 # correction factor
+    params['t_start'] = 0.0
+
+    params['x_v_dash'] 			= [2*np.zeros(params['n_cars'])]
+
+    v0 = params['v0']
+    init_v = params['init_v']
+    T = params['T']
+    a = params['a']
+    b = params['b']
+    delta = params['delta']
+    s0 = params['s0']
+    end_of_track = params['end_of_track']
+    n_cars = params['n_cars']
+    total_time = params['total_time']
+    for i in range(200,300,10):
+        params['t_steps'] 			= i # number of timesteps
+        params['t_step'] = (params['total_time']-params['t_start'])/float(params['t_steps'])
+        t_steps = params['t_steps']
+        params['IDM_model_num'] = 1
+
+        # Assign initial velocities (30m/s)
+        v = np.ones(n_cars) * init_v
+        # Assign initial positions
+        x_vec = np.linspace(0,end_of_track-end_of_track/6,n_cars)
+        # reverse positions so that car 0 is leading
+        x_vec = x_vec[::-1]
+        # create 1D vector of positions followed by velocities
[…]

### 2016-12-09 code/HDM.py
removing junk file

@@ -1,72 +1,0 @@
-#https://en.wikipedia.org/wiki/Intelligent_driver_model
-
-
-import numpy as np
-import scipy as sp
-import scipy.integrate
-import matplotlib
-import matplotlib.pyplot as plt
-import pdb
-
-v0 = 30.
-T = 1.5
-a = 1.
-b = 3.
-delta = 4.
-s0 = 2.
-end_of_track = 600
-t_steps = 100000
-
-
-def x_dash(t, v):
-	return v
-
-def x(x0, v, t):
-	return x0 + v*t
-
-def s_star(v, delta_v):
-	return s0 + v*T + (v*delta_v/(2*np.sqrt(a*b)))**2
-
-def x_v_dash(x_v, t):
-	# pdb.set_trace()
-	x_v = x_v.reshape(2,-2)
-	v = x_v[1,:]
-	x_vec = x_v[0,:]
-	delta_v = v.copy()
-
-	s = x_vec.copy()
-
-	for i in range(len(s)):
-		if i == len(s)-1:
-			delta_v[i] = v[0] - v[-1]
-			s[i] =  -end_of_track%(x_vec[0] - x_vec[-1])
-		else:
-			delta_v[i] =  v[i] - v[i+1]
-			s[i] =  x_vec[i] - x_vec[i+1]
-	# print x_vec
-	# print s
-	# print delta_v
-	# if (x_vec[-1] < 0.1):
-	# pdb.set_trace()
-	x_v = np.concatenate(([v], [a*(1-(v/v0)**delta - (s_star(v, delta_v)/s)**2)]), axis=0).reshape(1,-1)[0]
-	return x_v
-
-v = np.ones(50) * 30
-x_vec = np.linspace(0,490,50)
-x_v_vec = np.concatenate(([x_vec], [v]), axis=0).reshape(1,-1)[0]
-
-ts = np.linspace(0,1000,t_steps)
-
-y_s = sp.integrate.odeint(x_v_dash, y0=x_v_vec, t=ts)
-
-fig, ax = plt.subplots(1,1, figsize=(8,8))
-# pdb.set_trace()q
-y_s = y_s.reshape(t_steps,-2,2)
-# for car in range(50):
-# 	ax.plot(ts, x_vec)
-for car in range(50):
-	# pdb.set_trace()
-	print y_s[:,car,0]
-	ax.plot(ts, y_s[:,car,0])
-# print xs
-plt.show()

### 2016-12-09 code/optimize_driver.py
starting code for driver optimization

@@ -1,0 +1,322 @@
+import numpy as np
+import scipy as sp
+import scipy.integrate
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.animation as animation
+import pdb
+from intelligent_driver_model import *
+from human_driver_model import *
+
+def runge_kutta_4_driverparams(x_v_vec_k, x_v_dash, t_k, h, params, past, driveridx, driverparams):
+	k1=x_v_dash(x_v_vec_k,t_k, params,driveridx, driverparams)
+	k2=x_v_dash(x_v_vec_k+.5*h*k1,t_k+.5*h, params,driveridx, driverparams)
+	k3=x_v_dash(x_v_vec_k+.5*h*k2,t_k+.5*h, params,driveridx, driverparams)
+	k4=x_v_dash(x_v_vec_k+h*k3,t_k+h, params,driveridx, driverparams)
+	x_v_vec_k_next = x_v_vec_k + h/6. * (k1 + 2*k2 + 2*k3 + k4)
+	# store the derivative at each timestep
+	params['x_v_dash'].append(k1)
+	return x_v_vec_k_next
+
+def x_v_dash_driver(x_v, t, params, driveridx, driverparams):
+	# Compute derivative of position and velocity
+	x_v = x_v.reshape(2,-2)
+	# get velocities
+	v = x_v[1,:]
+	# get positions
+	x_vec = x_v[0,:]
+	## Follow the leader ##
+	# Note: Car i follows car i-1
+	# Since this is a ring track, car 0 follows car n-1
+	# for vehicle i, delta_v = v[i] - v[i-1]
+	vl = np.roll(v,1)
+	# vl[0] = float('inf')
+	# Compute difference in speeds between current car and leading car
+	delta_v = v - vl
+
+	# Compute gap
+	# for vehicle i, s = x_vec[i-1] - x_vec[i]
+	# put s for car zero within the bounds of the track
+	s = np.roll(x_vec,1) - x_vec
+	s[0] += params['end_of_track']
+
+	# s[0] = float('inf')
+	# Compute acceleration
+	if params['IDM_model_num'] == 0:
+		# Standard IDM
+		dvdt = a_IDM(v,s,delta_v,params)
+        dvdt_2 = a_IDM(v,s,delta_v,driverparams)
+        dvdt[driveridx] = dvdt_2[driveridx]
+	else:
+		# compute Improved IDM acceration function
[…]

### 2016-12-09 code/optimize_driver.py
starting code for driver optimization

@@ -204,6 +204,64 @@ def x_v_dash2_driver_human(x_v, t,params,past,driveridx, driverparams):
 	past['past_dvdt'].append(dvdt)
 	return x_v
 
+# define a wrapper function to determine what cost is being calculated
+# def cost_wrapper(optimize_params, cost_func, optimize_variables, params):
+# 	for var, param in zip(optimize_variables, optimize_params):
+# 		params[var] = param
+#
+# 	simulation_result = intelligent_driver_model.run_simulation(params)
+# 	# pdb.set_trace()
+# 	x_v_dash = np.array(params['x_v_dash'])
+#
+# 	displacement = simulation_result[:,:params['n_cars']].T
+# 	velocity = simulation_result[:,params['n_cars']:].T
+# 	acceleration = x_v_dash[:,params['n_cars']:].T
+#
+# 	# We should be able to write cost functions based on these three variables
+# 	c = cost_func(displacement, velocity, acceleration, params)
+# 	return c
+#
+# # Define the cost functions:
+#
+# # Cost Function 1
+# # Minimizing the total travel time -> Page 419
+# # Note: Minimising the travel time is the same as maximising the
+# # displacement across the time interval
+# def maximise_displacement(displacement, velocity, acceleration, params):
+# 	return 1-np.sum(displacement[:,-1])/float(params['n_cars'])
+#
+# # Cost Function 2
+# # Maximizing the driving comfort (minimise discomfort) -> Page 419
+# def comfort_cost(displacement, velocity, acceleration, params):
+# 	tau0 = 	1. #s -> characteristic time
+# 	a0 = 		1. # m/s**2
+# 	T0 = 		1. # arbitrary normalisation constant
+#
+# 	# enforce smooth boundary condition
+# 	acceleration_dash = np.concatenate((np.zeros((params['n_cars'],1)), np.diff(acceleration)/params['t_step']), axis=1)
+# 	integrand = np.sum(acceleration**2 + tau0**2*acceleration_dash**2, axis=0)
+# 	cost = 1/(T0*a0**2)*sp.integrate.trapz(
+# 									y=integrand,
+# 									axis=0
[…]

### 2016-12-09 code/optimize_driver.py
running

@@ -7,8 +7,9 @@ import matplotlib.animation as animation
 import pdb
 from intelligent_driver_model import *
 from human_driver_model import *
+import copy
 
-def runge_kutta_4_driverparams(x_v_vec_k, x_v_dash, t_k, h, params, past, driveridx, driverparams):
+def runge_kutta_4_driverparams(x_v_vec_k, x_v_dash, t_k, h, params, driveridx, driverparams):
 	k1=x_v_dash(x_v_vec_k,t_k, params,driveridx, driverparams)
 	k2=x_v_dash(x_v_vec_k+.5*h*k1,t_k+.5*h, params,driveridx, driverparams)
 	k3=x_v_dash(x_v_vec_k+.5*h*k2,t_k+.5*h, params,driveridx, driverparams)
@@ -42,23 +43,25 @@ def x_v_dash_driver(x_v, t, params, driveridx, driverparams):
 
 	# s[0] = float('inf')
 	# Compute acceleration
+	# pdb.set_trace()
 	if params['IDM_model_num'] == 0:
 		# Standard IDM
 		dvdt = a_IDM(v,s,delta_v,params)
-        dvdt_2 = a_IDM(v,s,delta_v,driverparams)
-        dvdt[driveridx] = dvdt_2[driveridx]
+		# pdb.set_trace()
+		dvdt_2 = a_IDM(v,s,delta_v,driverparams)
+		dvdt[driveridx] = dvdt_2[driveridx]
 	else:
 		# compute Improved IDM acceration function
 		# Note: we compute this acceleration function for both IIDM and ACC
 		# For ACC, this serves as dvdt
 		dvdt = a_IIDM(v,s,delta_v,params)
-        dvdt_2 = a_IIDM(v,s,delta_v,driverparams)
-        dvdt[driveridx] = dvdt_2[driveridx]
+		dvdt_2 = a_IIDM(v,s,delta_v,driverparams)
+		dvdt[driveridx] = dvdt_2[driveridx]
 	if params['IDM_model_num'] == 2:
 		# AAC IDM
 		dvdt = a_ACC(s, v, vl, dvdt, params)
-        dvdt_2 = a_ACC(s, v, vl, dvdt,driverparams)
-        dvdt[driveridx] = dvdt_2[driveridx]
+		dvdt_2 = a_ACC(s, v, vl, dvdt,driverparams)
+		dvdt[driveridx] = dvdt_2[driveridx]
[…]

### 2016-12-09 code/optimize_driver.py
trying inf cost for instability

@@ -9,12 +9,19 @@ from intelligent_driver_model import *
 from human_driver_model import *
 import copy
 
-def runge_kutta_4_driverparams(x_v_vec_k, x_v_dash, t_k, h, params, driveridx, driverparams):
-	k1=x_v_dash(x_v_vec_k,t_k, params,driveridx, driverparams)
-	k2=x_v_dash(x_v_vec_k+.5*h*k1,t_k+.5*h, params,driveridx, driverparams)
-	k3=x_v_dash(x_v_vec_k+.5*h*k2,t_k+.5*h, params,driveridx, driverparams)
-	k4=x_v_dash(x_v_vec_k+h*k3,t_k+h, params,driveridx, driverparams)
-	x_v_vec_k_next = x_v_vec_k + h/6. * (k1 + 2*k2 + 2*k3 + k4)
+def runge_kutta_4_driverparams(x_v_vec_k, x_v_dash, t_k, h, params, driveridx, driverparams, past):
+	if x_v_dash == x_v_dash_driver:
+		k1=x_v_dash(x_v_vec_k,t_k, params,driveridx, driverparams)
+		k2=x_v_dash(x_v_vec_k+.5*h*k1,t_k+.5*h, params,driveridx, driverparams)
+		k3=x_v_dash(x_v_vec_k+.5*h*k2,t_k+.5*h, params,driveridx, driverparams)
+		k4=x_v_dash(x_v_vec_k+h*k3,t_k+h, params,driveridx, driverparams)
+		x_v_vec_k_next = x_v_vec_k + h/6. * (k1 + 2*k2 + 2*k3 + k4)
+	else:
+		k1=x_v_dash(x_v_vec_k,t_k, params,driveridx, driverparams, past)
+		k2=x_v_dash(x_v_vec_k+.5*h*k1,t_k+.5*h, params,driveridx, driverparams, past)
+		k3=x_v_dash(x_v_vec_k+.5*h*k2,t_k+.5*h, params,driveridx, driverparams, past)
+		k4=x_v_dash(x_v_vec_k+h*k3,t_k+h, params,driveridx, driverparams, past)
+		x_v_vec_k_next = x_v_vec_k + h/6. * (k1 + 2*k2 + 2*k3 + k4)
 	# store the derivative at each timestep
 	params['x_v_dash'].append(k1)
 	return x_v_vec_k_next
@@ -44,12 +51,15 @@ def x_v_dash_driver(x_v, t, params, driveridx, driverparams):
[…]

### 2016-12-09 code/optimize_driver.py
tried scipy brute, no good

@@ -261,10 +261,14 @@ def run_hdm_simulation(params, driveridx, driverparams, past):
 
 
 # define a wrapper function to determine what cost is being calculated
-def cost_wrapper(optimize_params, cost_func, optimize_variables, params, driveridx, driverparams, past):
-	for var, param in zip(optimize_variables, optimize_params):
-		driverparams[var] = param
-
+# def cost_wrapper(optimize_params, cost_func, optimize_variables, params, driveridx, driverparams, past):
+def cost_wrapper(z, *costparams):
+	s0, T = z
+	cost_func, params, driveridx, driverparams, past = costparams
+	# for var, param in zip(optimize_variables, optimize_params):
+	# 	driverparams[var] = param
+	driverparams['s0'] = s0
+	driverparams['T'] = T
 	# simulation_result = run_idm_simulation(params, driveridx, driverparams)
 	simulation_result = run_hdm_simulation(params, driveridx, driverparams, past)
 
@@ -279,6 +283,7 @@ def cost_wrapper(optimize_params, cost_func, optimize_variables, params, driveri
 	c = cost_func(displacement, velocity, acceleration, params, driveridx)
 	# pdb.set_trace()
 	if sum(sum(np.isnan(np.array(x_v_dash))))>0:
+		print ('nan')
 		c = inf
 	return c
 #
@@ -344,16 +349,16 @@ if __name__ == '__main__':
 	params['delta'] = 4.0 # Acceleration exponent
 	params['s0'] = 10.0 # minimum gap (in m)
 	params['end_of_track'] = 800 # in m
-	params['t_steps'] = 10000 # number of timesteps
+	params['t_steps'] = 100 # number of timesteps
 	params['t_start'] = 0.0
 	params['n_cars'] = 10 # number of vehicles
-	params['total_time'] = 3000 # total time (in s)
+	params['total_time'] = 30 # total time (in s)
 	params['c'] = 0.99 # correction factor
 	params['t_step'] = (params['total_time'] - params['t_start'])/params['t_steps']
 	params['Tr'] = .6 # Reaction time
 	params['Vs'] = 0.1 # Variation coefficient of gap estimation error
 	params['sigma_r'] = 0.01 # estimation error for the inverse TTC
-	params['sigma_a']= 1.  # magnitude of acceleration noise
[…]

### 2016-12-10 code/optimize_driver.py
now optimizing by brute force

@@ -261,14 +261,14 @@ def run_hdm_simulation(params, driveridx, driverparams, past):
 
 
 # define a wrapper function to determine what cost is being calculated
-# def cost_wrapper(optimize_params, cost_func, optimize_variables, params, driveridx, driverparams, past):
-def cost_wrapper(z, *costparams):
-	s0, T = z
-	cost_func, params, driveridx, driverparams, past = costparams
-	# for var, param in zip(optimize_variables, optimize_params):
-	# 	driverparams[var] = param
-	driverparams['s0'] = s0
-	driverparams['T'] = T
+def cost_wrapper(optimize_params, cost_func, optimize_variables, params, driveridx, driverparams, past):
+# def cost_wrapper(z, *costparams):
+	# s0, T = z
+	# cost_func, params, driveridx, driverparams, past = costparams
+	for var, param in zip(optimize_variables, optimize_params):
+		driverparams[var] = param
+	# driverparams['s0'] = s0
+	# driverparams['T'] = T
 	# simulation_result = run_idm_simulation(params, driveridx, driverparams)
 	simulation_result = run_hdm_simulation(params, driveridx, driverparams, past)
 
@@ -284,7 +284,7 @@ def cost_wrapper(z, *costparams):
 	# pdb.set_trace()
 	if sum(sum(np.isnan(np.array(x_v_dash))))>0:
 		print ('nan')
-		c = inf
+		c = np.inf
 	return c
 #
 # # Define the cost functions:
@@ -294,7 +294,7 @@ def cost_wrapper(z, *costparams):
 # # Note: Minimising the travel time is the same as maximising the
 # # displacement across the time interval
 def maximise_displacement(displacement, velocity, acceleration, params,driveridx):
-	return 1-displacement[driveridx,-1]
+	return displacement[driveridx,-1], np.sum(displacement[:,-1]/params['n_cars'])
 #
 # # Cost Function 2
 # # Maximizing the driving comfort (minimise discomfort) -> Page 419
@@ -349,10 +349,10 @@ if __name__ == '__main__':
 	params['delta'] = 4.0 # Acceleration exponent
 	params['s0'] = 10.0 # minimum gap (in m)
 	params['end_of_track'] = 800 # in m
-	params['t_steps'] = 100 # number of timesteps
+	params['t_steps'] = 2000 # number of timesteps
 	params['t_start'] = 0.0
[…]

### 2016-12-10 code/optimize_driver.py
returning inf cost for crashes

@@ -283,8 +283,11 @@ def cost_wrapper(optimize_params, cost_func, optimize_variables, params, driveri
 	c = cost_func(displacement, velocity, acceleration, params, driveridx)
 	# pdb.set_trace()
 	if sum(sum(np.isnan(np.array(x_v_dash))))>0:
-		print ('nan')
-		c = np.inf
+		if np.isnan(x_v_dash[-1][driveridx])==False:
+			print ('nan')
+			c = displacement[driveridx,-1], np.inf
+		else:
+			c = np.inf, np.inf
 	return c
 #
 # # Define the cost functions:
@@ -294,7 +297,7 @@ def cost_wrapper(optimize_params, cost_func, optimize_variables, params, driveri
 # # Note: Minimising the travel time is the same as maximising the
 # # displacement across the time interval
 def maximise_displacement(displacement, velocity, acceleration, params,driveridx):
-	return displacement[driveridx,-1], np.sum(displacement[:,-1]/params['n_cars'])
+	return displacement[driveridx,-1], np.sum(np.std(velocity, axis=0))
 #
 # # Cost Function 2
 # # Maximizing the driving comfort (minimise discomfort) -> Page 419
@@ -351,14 +354,14 @@ if __name__ == '__main__':
 	params['end_of_track'] = 800 # in m
 	params['t_steps'] = 2000 # number of timesteps
 	params['t_start'] = 0.0
-	params['n_cars'] = 10 # number of vehicles
+	params['n_cars'] = 50 # number of vehicles
 	params['total_time'] = 600 # total time (in s)
 	params['c'] = 0.99 # correction factor
 	params['t_step'] = (params['total_time'] - params['t_start'])/params['t_steps']
 	params['Tr'] = .6 # Reaction time
 	params['Vs'] = 0.1 # Variation coefficient of gap estimation error
 	params['sigma_r'] = 0.01 # estimation error for the inverse TTC
-	params['sigma_a']= .01  # magnitude of acceleration noise
+	params['sigma_a']= .1  # magnitude of acceleration noise
 	params['tau_tilde'] = 20.0 # persistence time of estimation errors (in s)
 	params['tau_a_tilde'] =  1.0 # persistence time of acceleration noise (in s)
 	v0 = params['v0']

### 2016-12-10 discussion.tex
added formulas to paper

@@
 \section{Introduction}
-The modeling of traffic at traffic lights, on highways, in dense streams etc.
+\paragraph{}The modeling of traffic at traffic lights, on highways, in dense streams etc.
 can often present a problem that is too complicated to solve analytically.
@@
 \section{Macroscopic Perspective of Traffic Interaction}
-There are generally two ways to model traffic.
+\paragraph{}There are generally two ways to model traffic.
 In the macroscopic perspective, traffic is viewed as a fluid or gas with a given maximum density moving according to the laws of conservation of mass.
@@
 
-\section{Microscopic Perspective of Traffic Interaction - The Analytical Solution}
-
 \section{Microscopic Perspective of Traffic Interaction : \\Intelligent Driver Models}
-In the microscopic perspective, individual cars are modeled as particles that move according to a relationship to their leading particle(s).
+\paragraph{}In the microscopic perspective, individual cars are modeled as particles that move according to a relationship to their leading particle(s).
 For this section of the problem, we will model the behaviors of cars in a stream of traffic individually.
@@
 Furthermore, we will use different controllers to simulate the means by which cars can control their acceleration and will be able to simulate the effect that autonomous cars may have for the throughput of a stream of traffic for varying car densities.
-This numerical analysis can be contrasted to the analytical solution that was presented in (**another section 2?**).\par
-A car`s speed is governed by the laws of motion:
-\begin{equation}
-v_f = v_i + a\Delta t
-\end{equation}
 
-An individual car can be modeled as such over a 100,000m track and asked to travel at $v_{target} = 60miles.h^{-1}$ (i.e.
+$$\dot{v} = a \left[1 - \left(\frac{v}{v_0}\right)^{\delta} - \left(\frac{s*(v,\Delta v)}{s}\right)^{2}\right]$$
+
+$$s*(v, \Delta v) = s_0 + max\left(0, vT + \frac{v \Delta v}{2 \sqrt{ab}} \right)$$
+
+$$a_{free}(v)= \begin{cases}
+a \left[ 1 - (\frac{v}{v_0})^\delta \right] & v \le v_0\\
+-b \left[ 1 - (\frac{v_0}{v})^{a\delta/b} \right]& v > v_0
+\end{cases}$$
+
+$$\frac{dv}{dt}\Bigr|_{v\le v_0}= \begin{cases}
+a (1-z^2) & z = \frac{s*(v, \Delta v}{s} \ge 1\\
+a_{free}(1 - z^{(2a)/a_{free}})& otherwise
+\end{cases}$$
+
+$$\frac{dv}{dt}\Bigr|_{v> v_0}= \begin{cases}
+a_{free} + a (1-z^2) & z(v, \Delta v) \ge 1\\
+a_{free} & otherwise
+\end{cases}$$
+
+$$a_{CAH}(s,v,v_l, \dot{v}_l)= \begin{cases}
+\frac{v^2\tilde{a}_l}{v_l^2 - 2 \tilde(a)_l} & v_l(v-v_l) \le -2s\tilde{a}_l\\
+\tilde{a}_l - \frac{(v-v_l)^2 \Theta (v-v_l)}{2s} & otherwise
+\end{cases}$$
+
+$$\tilde{a}_l(\dot(v)_l) = min(\dot{v}_l, a)$$
+
+$$a_{ACC}= \begin{cases}
+a_{IIDM} & a_{IIDM} \ge a_{CAH}\\
+(1-c)a_{IIDM} + c\left[a_{CAH} + b tanh(\frac{a_{IIDM}-a_{CAH}}{b}) \right] & otherwise
+\end{cases}$$
+
+%This numerical analysis can be contrasted to the analytical solution that was presented in (**another section 2?**).\par
+%A car`s speed is governed by the laws of motion:
+%\begin{equation}
+%v_f = v_i + a\Delta t
+%\end{equation}
+%
+%An individual car can be modeled as such over a 100,000m track and asked to travel at $v_{target} = 60miles.h^{-1}$ (i.e.
 $26.8m.s^{-1}$).
@@
 We can therefore model his fluctuations as random acceleration with the corresponding adjustments to correct the acceleration:
-\begin{equation}
-a_{correction_i} = \alpha(v_{instant} - v_{target}) + N(0,\sigma_1)
-\end{equation}
-
-We note that for more than one car in a stream of traffic the car's acceleration is then also dependent on the car ahead of it:
-\begin{equation}\label{eq:interaction}
-a_{interaction_i} = \beta (s_{instant} - s_{target})
-\end{equation}
-
-In equation \ref{eq:interaction}, we see that the acceleration of the car is also dependent on the distance between any one car and the car that is directly ahead of it.
+%\begin{equation}
+%a_{correction_i} = \alpha(v_{instant} - v_{target}) + N(0,\sigma_1)
+%\end{equation}
+%
+%We note that for more than one car in a stream of traffic the car's acceleration is then also dependent on the car ahead of it:
+%\begin{equation}\label{eq:interaction}
+%a_{interaction_i} = \beta (s_{instant} - s_{target})
+%\end{equation}
+%
+%In equation \ref{eq:interaction}, we see that the acceleration of the car is also dependent on the distance between any one car and the car that is directly ahead of it.
 The final acceleration of the car is therefore:
-\begin{equation}\label{eq:acceleration}
-a_{i} = a_{correction_i} + a_{interaction_i}
-\end{equation}
-
-If we assume the cars should travel at $v_{target}$ and that they should maintain an $s_{target} = 5m$ spacing between them, we can then model the velocity of the car as:
-
-\begin{equation}\label{eq:velocity}
-v_{i} = \frac{d(a_i)}{dt}
-\end{equation}
+%\begin{equation}\label{eq:acceleration}
+%a_{i} = a_{correction_i} + a_{interaction_i}
+%\end{equation}
+%
+%If we assume the cars should travel at $v_{target}$ and that they should maintain an $s_{target} = 5m$ spacing between them, we can then model the velocity of the car as:
+%
+%\begin{equation}\label{eq:velocity}
+%v_{i} = \frac{d(a_i)}{dt}
+%\end{equation}
 
@@
 
-\section{Stability Comparison of a Variety of Numerical Integration Schemes}
+\paragraph{}To generate numerical simulations of Traffic Systems, we first consider the case of 50 cars on a circular track.
+We explored scenarios of all IDM cars and all HDM cars.
+Under these conditions, we see the IDM models generate more `phantom` traffic waves than the HDM models, especially for larger numbers of HDM car lookahead.
+This is because the human drivers tend to appear more conservative in a situation like this, with no external perturbations, as the multi-car lookahead prevents them from getting overly close to the car directly in front of them.
+
+% \begin{figure}[H]
+% \centering
+% \includegraphics[keepaspectratio=true,scale=0.6]{problem3d}
+% \caption{Plot of temperature cross section in pipe at various time-steps of simulation}\label{visina8}
+% \end{figure}
+
+% \begin{figure}[H]
+% \centering
+% \includegraphics[keepaspectratio=true,scale=0.6]{problem3d}
+% \caption{Plot of temperature cross section in pipe at various time-steps of simulation}\label{visina8}
+% \end{figure}
+
+\paragraph{}We then moved on to simulations of a platoon of cars with some cars being controlled by IDM-type models (in particular, Adaptive Cruise Control) and some cars being controlled by multi-car lookahead HDM models.
+After allowing the models to run for some time to develop their natural spacing, we then perturbed the acceleration of the lead car, causing it to brake and then resume again.
+We then observe both the magnitude and length of the volatility of the velocities in the following cars after the event.
+We find that in general, as one would expect, a larger proportion of IDM-controlled vehicles results in a quicker dissipation of the perturbation as well as a lowered probability of a vehicle collision.
+\textbf{RESULTS HERE}
+
+\section{Stability and Error Comparison of a Variety of Numerical Integration Schemes}
+Because there is no analytical solution given for our highly-nonlinear problem, to analyze the error of our numerical integration schemes, we compared to the python \texttt{odeint} solver function.
+For large enough time steps, this shows us error decreasing on a log scale for each of our higher order methods.
+However, for small time steps, the `error` for 4th and 5th order methods converges, as the \texttt{odeint} solver presumably decides that 4th order is sufficiently accurate at these time steps and stops using higher order methods.
+\textbf{RESULTS HERE}
+For our purposes, we desired to use higher order methods such that we could use larger timesteps and hopefully decrease running time of our programs while still maintaining stability in the differential equation systems.
+To display the effectiveness of these higher order methods, we ran one of our simple models for a variety of time steps over the same total time with the different solvers and noted the number of time steps at which each method stabilized.
+\textbf{RESULTS HERE}
 
@@
 
+%as a function of the reaction time Tr and the attention span ?t depends strongly on
+%the number of considered vehicles. When considering na = 5 leaders, the critical
+%effective reaction time Tr + ?t/2 at the stability limit is about twice as large as
+%the corresponding critical value without multi-anticipation (na = 1). Particularly,
+%traffic can be stable even if the reaction time exceeds the average time headway. This
+%agrees with everyday observations but cannot be realized in simulations without
+%multi-anticipation. There are limits, however: Anticipating more than five vehicles
+%ahead will change the dynamics insignificantly
+
 \section{Optimization of Traffic Systems}
-
+\paragraph{}We set up a number of optimization programs to attempt to search for the optimal values of some of the many parameters for the system as a whole and for individual drivers for a number of cost functions, including maximizing displacement and maximizing comfort (minimizing jerk).
+The initial trials for these systems were run on IDM models, and the results are largely what one would suspect.
+When attempting to maximize total displacement, the cars drive quickly and close together, resulting in the creation of traffic waves in the system.
+When attempting to minimize jerk, the cars drive much more conservatively, as one would expect, never coming within the range of another car that would cause them to change acceleration.
+\textbf{RESULTS HERE}
+\paragraph{}We then attempted to optimize behavior for a single driver and were interested to see what implications this would have for the system as a whole.
+However, unsurprisingly, the following behaviors in the IDM are such that a single driver can tailgate and maximize target speed basically as much as we will allow without ever disrupting the rest of the system.
+Even when adding some stochasticity to the lead car, we were unable to see the rest of the system suffer from traffic waves due to the behavior of the second car, because that behavior would simply be quickly dissipated by the third car.
+In fact, it seems to be true generally of these systems that unless the aggressive driving behavior of a car causes other cars to collide behind it, the greater distance achieved by one car inevitably results in other cars also achieving a greater distance/average velocity.
+However, this is likely to occur with much greater volatility in the velocity of the cars, i.e.
+a less pleasant ride.
+Then, to quantify how the erratic driving of a single car may result in penalties to the system as a whole but not to that car, we must analyze the cost to the other cars in terms of the volatility of their ride speed and possible collisions.\\
+\textbf{RESULTS HERE}
+\paragraph{}Next, we attempted to replicate these experiments with the more complicated Human Driver Model.
+However, due to the highly piecewise nature of the functions we are attempting to optimize over, we were unable to use traditional minimization solvers.
+Instead, we discretized the parameter space we were interested in exploring and searched for the set of parameters which would minimize our target cost function.
+\textbf{RESULTS HERE}
 \section{TODO}
-run stability analysis of Runge Kutta schemes \\
-develop a way to run smaller timesteps over any portion we think is likely to explode (although not actually sure how much this will help) \\
 develop a cost function for emissions/energy usage and optimize over that \\

### 2016-12-10 code/optimize_driver.py

@@ -141,7 +141,7 @@ def x_v_dash2_driver_human(x_v, t,params,driveridx, driverparams,past):
 	# for vehicle i, s = x_vec[i-1] - x_vec[i]
 	s = np.roll(x_vec,1) - x_vec
 	# put s for car zero within the bounds of the track
-	s[0] += end_of_track
+	s[0] = 1000
 	# Compute index of this timestep
 	index = int(math.floor(t/t_step)) if math.floor(t/t_step) < t_steps else t_steps-1
 	# compute estimate of gap
@@ -179,20 +179,40 @@ def x_v_dash2_driver_human(x_v, t,params,driveridx, driverparams,past):
 		dvdt = np.zeros(v.shape)
 		# calculate the acceleration of each vehicle one at a time
 		# As written in Eq 12.20 of the textbook
-		for alpha in xrange(n_cars):
+		dvdt[0] = free_term[0]
+		for alpha in xrange(1,n_cars):
 			if alpha == driveridx:
 				free_term_alpha = free_term[alpha]
 				int_term = 0.0
 				v_alpha_prog = np.array([v_prog[alpha]])
-				for beta in xrange(alpha-n_a,alpha):
-					# we want to sum from beta+1 to alpha
-					# precompute these indices to avoid slicing issues
-					sum_idxs = np.arange(beta+1,alpha+1)
-					#print sum_idxs
-					s_alpha_beta_prog = np.array([np.sum(s_prog[sum_idxs])])
-					#print s_alpha_beta_prog
-					v_beta_prog = np.array([v_l_prog[beta]])
-					int_term += a_IDM_int(v_alpha_prog,s_alpha_beta_prog,v_alpha_prog-v_beta_prog,driverparams)
+				if alpha==1:
+					for beta in xrange(0,1):
+						sum_idxs = np.arange(beta+1,alpha+1)
+						#print sum_idxs
+						s_alpha_beta_prog = np.array([np.sum(s_prog[sum_idxs])])
+						#print s_alpha_beta_prog
+						v_beta_prog = np.array([v_l_prog[beta]])
+						int_term += a_IDM_int(v_alpha_prog,s_alpha_beta_prog,v_alpha_prog-v_beta_prog,driverparams)
+				elif alpha==2:
+					for beta in xrange(0,2):
+						for beta in xrange(0,1):
+							sum_idxs = np.arange(beta+1,alpha+1)
+							#print sum_idxs
+							s_alpha_beta_prog = np.array([np.sum(s_prog[sum_idxs])])
[…]

### 2016-12-10 discussion.tex
added formulas to paper

@@
 We then observe both the magnitude and length of the volatility of the velocities in the following cars after the event.
-We find that in general, as one would expect, a larger proportion of IDM-controlled vehicles results in a quicker dissipation of the perturbation as well as a lowered probability of a vehicle collision.
+We find that in general, as one would expect, a larger proportion of IDM-controlled vehicles results in a quicker dissipation of the perturbation as well as a lowered probability of a vehicle collision.\\
 \textbf{RESULTS HERE}
@@
 For large enough time steps, this shows us error decreasing on a log scale for each of our higher order methods.
-However, for small time steps, the `error` for 4th and 5th order methods converges, as the \texttt{odeint} solver presumably decides that 4th order is sufficiently accurate at these time steps and stops using higher order methods.
+However, for small time steps, the `error` for 4th and 5th order methods converges, as the \texttt{odeint} solver presumably decides that 4th order is sufficiently accurate at these time steps and stops using higher order methods.\\
 \textbf{RESULTS HERE}
-For our purposes, we desired to use higher order methods such that we could use larger timesteps and hopefully decrease running time of our programs while still maintaining stability in the differential equation systems.
-To display the effectiveness of these higher order methods, we ran one of our simple models for a variety of time steps over the same total time with the different solvers and noted the number of time steps at which each method stabilized.
+\paragraph{}For our purposes, we desired to use higher order methods such that we could use larger timesteps and hopefully decrease running time of our programs while still maintaining stability in the differential equation systems.
+To display the effectiveness of these higher order methods, we ran one of our simple models for a variety of time steps over the same total time with the different solvers and noted the number of time steps at which each method stabilized.\\
 \textbf{RESULTS HERE}
@@
 When attempting to minimize jerk, the cars drive much more conservatively, as one would expect, never coming within the range of another car that would cause them to change acceleration.
+\\
 \textbf{RESULTS HERE}
@@
 Instead, we discretized the parameter space we were interested in exploring and searched for the set of parameters which would minimize our target cost function.
+\\
 \textbf{RESULTS HERE}

### 2016-12-10 project_paper/paper.tex

--- 2016-12-10 discussion.tex
@@
-\documentclass[11pt]{article}
-\usepackage{fullpage,amsmath,amsfonts,mathpazo,microtype,nicefrac,graphicx,verbatimbox,listings,hyperref,enumitem,amssymb,float,pdfpages}
-% Macro definitions
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\Q}{\mathbb{Q}}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\B}{\mathbb{B}}
-\newcommand{\p}{\partial}
-\newcommand{\Trans}{\mathsf{T}}
-\renewcommand{\vec}[1]{\mathbf{#1}}
-\newcommand{\vx}{\vec{x}}
-\newcommand{\vb}{\vec{b}}
+% Use only LaTeX2e, calling the article.cls class and 12-point type.
 
-\DeclareMathOperator{\rank}{rank}
+\documentclass[12pt]{article}
 
-% Include support for including a pdf
+% Users of the {thebibliography} environment or BibTeX should use the
+% scicite.sty package, downloadable from *Science* at
+% www.sciencemag.org/about/authors/prep/TeX_help/ .
+% This package should properly format in-text
+% reference calls and reference-list numbers.
 
-% \begin{lstlisting}[frame=single]
-% \end{lstlisting}
+\usepackage{scicite}
 
-% \begin{figure}[H]
-% \centering
-% \includegraphics[keepaspectratio=true,scale=0.6]{problem3d}
-% \caption{Plot of temperature cross section in pipe at various time-steps of simulation}\label{visina8}
-% \end{figure}
+% Use times if you have the font installed; otherwise, comment out the
+% following line.
+
+\usepackage{times,amsmath}
+
+% The preamble here sets up a lot of new/revised commands and
+% environments.  It's annoying, but please do *not* try to strip these
+% out into a separate .sty file (which could lead to the loss of some
+% information when we convert the file to other formats).  Instead, keep
+% them in the preamble of your main LaTeX source file.
+
+% The following parameters seem to provide a reasonable page setup.
+
+\topmargin 0.0cm
+\oddsidemargin 0.2cm
+\textwidth 16cm
+\textheight 21cm
+\footskip 1.0cm
+
+%The next command sets up an environment for the abstract to your paper.
+
+\newenvironment{sciabstract}{%
+\begin{quote} \bf}
+{\end{quote}}
+
+% If your reference list includes text notes as well as references,
+% include the following line; otherwise, comment it out.
+
+\renewcommand\refname{References and Notes}
+
+% The following lines set up an environment for the last note in the
+% reference list, which commonly includes acknowledgments of funding,
+% help, etc.  It's intended for users of BibTeX or the {thebibliography}
+% environment.  Users who are hand-coding their references at the end
+% using a list environment such as {enumerate} can simply add another
+% item at the end, and it will be numbered automatically.
+
+\newcounter{lastnote}
+\newenvironment{scilastnote}{%
+\setcounter{lastnote}{\value{enumiv}}%
+\addtocounter{lastnote}{+1}%
+\begin{list}%
+{\arabic{lastnote}.}
+{\setlength{\leftmargin}{.22in}}
+{\setlength{\labelsep}{.5em}}}
+{\end{list}}
+
+\renewcommand{\baselinestretch}{1.3} % 1.5 line spacing
+\setlength{\parindent}{4em}
+\setlength{\parskip}{1em}
+
+% Include your paper's title here
+
+\title{AM205 Project - Traffic Flow Modelling\\
+\Large{Microscopic Models of Traffic, Comparing the Intelligent Driver Model (with Adaptive Cruise Control) with the Human Driver Model}}
+
+% Place the author information here.  Please hand-code the contact
+% information and notecalls; do *not* use \footnote commands.  Let the
+% author contact information appear immediately below the author names
+% as shown.  We would also prefer that you don't change the type-size
+% settings shown here.
+
+\author
+{Anna Hilgard, Sam Daulton and Nicholas Hoernle}
+
+% Include the date command, but leave its argument blank.
+
+\date{}
+
+%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%
 
@@
 
-\lstset{language=Python, basicstyle=\scriptsize} % set language as python
+% Make the title.
 
-\section*{AM205 Project - Traffic Flow Modelling; do autonomous cars help? (I don't know about this but we can change it)}
+\maketitle
 
 \section{Introduction}
-\paragraph{}The modeling of traffic at traffic lights, on highways, in dense streams etc.
-can often present a problem that is too complicated to solve analytically.
-This is mainly due to the inherent random movements of cars, the number of individually moving cars and the many scenarios that can be analyzed.
-These scenarios can be modeled numerically by applying the equations of motion to a system of vehicles and observing the impact of a number of outcomes.
-In this project, we will present an analytical solution to a number of simplistic scenarios involving traffic (cars starting from a red light, cars approaching a red light etc).
-We will then present the same scenario using a numerically calculated solution, modeling the cars using a differential setting where one car's motion is dependent on the car that is in front of it.
-We will then extend the scope of this analysis and introduce a number of varying factors into the model to analyze the motion of cars along a straight road, analyze the cars along a circular track and finally include autonomous cars with the human controlled vehicles.
+
+\paragraph{}
+The modeling of traffic at traffic lights, on highways, in dense traffic streams, with varying situations etc can often present a problem that is too complicated to solve analytically.
+This is mainly due to the inherent random movements of cars, the number of individually moving cars and the many and varying scenarios that can be analyzed.
+In this project, we present two main models, the `Intelligent Driver Model' and the `Human Driver Model' and numerically simulate a number of test cases allowing us to study the underlying dynamics of this system.
+We focus on modeling the cars using a differential setting where one car's motion is dependent on the car that is in front of it (and sometimes dependent on multiple cars in front of it).
+We will then extend the scope of this analysis and introduce a number of varying factors into the model to analyze the motion of cars along a straight road, analyze the cars along a circular track and finally include a blend of the intelligent driving cars (comparable to automated cruise control autonomous cars) and cars with the human controlled vehicles.
 These autonomous cars have more precise, and faster reacting velocity controllers and the inclusion of vehicles such as these in the model is expected to smooth the flow of traffic and ultimately increase the density of traffic flow.
@@
 
-$$a_{free}(v)= \begin{cases}
+$$
+a_{free}(v)= \begin{cases}
 	a \left[ 1 - (\frac{v}{v_0})^\delta \right] & v \le v_0\\
--b \left[ 1 - (\frac{v_0}{v})^{a\delta/b} \right]& v > v_0
-\end{cases}$$
+	-b \left[ 1 - (\frac{v_0}{v})^{a\delta/b} \right] & v > v_0
+\end{cases}
+$$
 
@@
 \end{cases}$$
+
+$$\dot{v}(t) = a_{mic}\left[s(t-T_r), v(t-T_r), v_l(t-T_r) \right]$$
+
+$$u(t-Tr) = ru_{i-j-1} + (1-r)u_{i-j},  \, \, j = int\left(\frac{T_r}{\Delta t} \right), \, \, r = \frac{T_r}{\Delta t} - j$$
+
+$$ln s^{est} - ln s = V_s w_s(t) $$
+
+$$v^{est} - v_l = -s \sigma_r w_l(t)$$
 
@@
 
+\section{Traffic Flow Simulation: Microscopic models}
+
+\section{Models and Methods}
+
+\section{Results}
+
+\section{Discussion}
+
+\section{Conclusion}
+
+% Your references go at the end of the main text, and before the
+% figures.  For this document we've used BibTeX, the .bib file
+% scibib.bib, and the .bst file Science.bst.  The package scicite.sty
+% was included to format the reference numbers according to *Science*
+% style.
+
+\bibliography{scibib}
+
+\bibliographystyle{Science}
+
 \end{document}

### 2016-12-10 project_paper/paper.tex
Paper updates

@@
 \usepackage{times,amsmath}
+\usepackage{empheq}
+\usepackage[theorems,skins]{tcolorbox} % for nice boxed eqs
+\newtcolorbox{mymathbox}[1][]{colback=white, sharp corners, #1}
+
+% for figs
+\usepackage{graphicx}
 
@@
 
-\section{Macroscopic Perspective of Traffic Interaction}
+\section{Macroscopic Perspective of Single Lane Traffic Interactions}
 \paragraph{}There are generally two ways to model traffic.
@@
 
-\section{Microscopic Perspective of Traffic Interaction : \\Intelligent Driver Models}
-\paragraph{}In the microscopic perspective, individual cars are modeled as particles that move according to a relationship to their leading particle(s).
-For this section of the problem, we will model the behaviors of cars in a stream of traffic individually.
-We will then simulate the movement of this stream of cars and demonstrate some well known traffic phenomenon (cars starting from a red traffic light will propagate forward in a wave formation, cars traveling along a straight road will oscillate into a traffic jam).
-Furthermore, we will use different controllers to simulate the means by which cars can control their acceleration and will be able to simulate the effect that autonomous cars may have for the throughput of a stream of traffic for varying car densities.
+\section{Microscopic Perspective of Single Lane Traffic Interactions}
+\paragraph{} In the microscopic perspective, individual cars are modeled as particles that move according to a relationship with the particle(s) in front of them.
+Time-continuous microscopic models, commonly referred to as car-following models, model the characteristics of individual vehicles such as acceleration, velocity, position using a systems of coupled ordinary differential equations (ODEs).
+Car-following models are constained by assumptions that mimic realistic driving behavior such as a vehicle wants to maintain a safe driving gap between itself and vehicle in front of it (commonly called the \textit{leading vehicle}), a vehicle has a certain desired velocity, or there is a maximum comfortable acceleration.
+In the section, we consider a few variations of the Intelligent Driver Model (IDM), as well as Human Driver Model (HDM).
+We chose the IDM over other car-following models such as Gipps Model because the IDM has direct extensions to modeling human driving behavior using the HDM.
+\subsection{Intelligent Driver Models}
+In this section, we model streams of semi-autonomous vehicles that display the following driving characteristics:
+\begin{itemize}
+  \item No estimation errors: each vehicle is aware of its own speed, the gap between itself the car in front of it, and the speed of the car in front of it.
+  \item Zero response time: there is no delay between observing a stimulus (i.e. change in gap between itself and vehicle ahead of it or change the vehicle ahead's speed) and responding to it.
+  \item Constant attention: the vehicle is constantly looking for changes in speed or gap
+  \item Each vehicle is only aware of itself and the vehicle directly in front of it.  I.e. it does not account for changes in speed or gap of a vehicle two cars ahead of it.
+  \item Each vehicle is not aware of vehicles behind it or in other lanes
+  \item Vehicles are not aware other indicators such as brake lights, horns, headlight flashes, etc.
+\end{itemize}
+\subsubsection{Intelligent Driver Model (IDM)}
+\begin{figure}
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{../figures/system_ppt_fig}}
+  \caption{Snapshot of two cars in a system modeled using a car-following model.
+The leading car with respect to vehicle $\alpha$  is referred to as vehicle $l=\alpha-1$.
+Note that vehicles are represented as particles, and therefore, the vehicles themselves have negligible length.}
+\end{figure}
+The IDM is a simple, accident-free model that can model traffic systems under all realistic traffic conditions.
+The acceleration function, shown in Eq (1), of a particular vehicle $\alpha$ is a function the current gap between itself and the leading vehicle ($s_\alpha$), the desired speed $v_0$, and the speed of the vehicle leading vehicle $v_l$.
+The first term in Eq (1) is the \textit{free acceleration}\---the acceleration that a vehicle would take if on an open road, without considerations of desired speed or gap.
+The second term in Eq (1) is the effect on the acceleration of the vehicle approaching its desired speed.
+The third term in Eq(1) is the effect on the acceleration of relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance.
+The desired gap distance given by Eq (4) the sum of two terms the steady state desired safe distance $s_0+v_\alpha T$ and a dynamical term $\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$ representing the intelligent vehicle's braking strategy.
 
-$$\dot{v} = a \left[1 - \left(\frac{v}{v_0}\right)^{\delta} - \left(\frac{s*(v,\Delta v)}{s}\right)^{2}\right]$$
+\begin{mymathbox}[title=Parameters, colframe=blue!30!black]
+  \begin{itemize}
+    \item $a$: maximum acceleration
+    \item $b$: maximum comfortable deceleration
+    \item $s_0$: minimum bumper-to-bumper gap (between vehicle and leading vehicle)
+    \item $v_0$: desired speed
+    \item $T$: minimum time gap (between vehicle and leading vehicle)
+    \item $\delta$: acceleration exponent
+  \end{itemize}
+\end{mymathbox}
 
-$$s*(v, \Delta v) = s_0 + max\left(0, vT + \frac{v \Delta v}{2 \sqrt{ab}} \right)$$
+\begin{mymathbox}[ams gather, title=IDM Governing Equations, colframe=blue!30!black]
+  \dot{v}_\alpha = a \left[1 - \left(\frac{v_\alpha}{v_0}\right)^{\delta} - \left(\frac{s^*(v_\alpha,\Delta v_\alpha)}{s_\alpha}\right)^{2}\right]\\
+  s_\alpha = x_{\alpha-1}-x_\alpha=x_{l}-x_\alpha\\
+  \Delta v_\alpha =v_\alpha-v_{\alpha-1}=v_\alpha-v_l\\
+  s^*(v_\alpha, \Delta v_\alpha) = s_0 + \max\left(0, v_\alpha T + \frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}} \right)
+\end{mymathbox}
 
+\subsubsection{Improved Intelligent Driver Model (IIDM)}
+
+\subsubsection{Adaptive Cruise Control Model (ACC)}
+
+\paragraph{}
 $$

### 2016-12-10 project_paper/paper.tex
Added IIDM Section

@@
 \end{figure}
+\paragraph{}
 The IDM is a simple, accident-free model that can model traffic systems under all realistic traffic conditions.
 The acceleration function, shown in Eq (1), of a particular vehicle $\alpha$ is a function the current gap between itself and the leading vehicle ($s_\alpha$), the desired speed $v_0$, and the speed of the vehicle leading vehicle $v_l$.
-The first term in Eq (1) is the \textit{free acceleration}\---the acceleration that a vehicle would take if on an open road, without considerations of desired speed or gap.
-The second term in Eq (1) is the effect on the acceleration of the vehicle approaching its desired speed.
-The third term in Eq(1) is the effect on the acceleration of relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance.
-The desired gap distance given by Eq (4) the sum of two terms the steady state desired safe distance $s_0+v_\alpha T$ and a dynamical term $\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$ representing the intelligent vehicle's braking strategy.
 
-\begin{mymathbox}[title=Parameters, colframe=blue!30!black]
+\begin{mymathbox}[title=IDM Parameters, colframe=blue!30!black]
   \begin{itemize}
@@
 \begin{mymathbox}[ams gather, title=IDM Governing Equations, colframe=blue!30!black]
-  \dot{v}_\alpha = a \left[1 - \left(\frac{v_\alpha}{v_0}\right)^{\delta} - \left(\frac{s^*(v_\alpha,\Delta v_\alpha)}{s_\alpha}\right)^{2}\right]\\
-  s_\alpha = x_{\alpha-1}-x_\alpha=x_{l}-x_\alpha\\
-  \Delta v_\alpha =v_\alpha-v_{\alpha-1}=v_\alpha-v_l\\
-  s^*(v_\alpha, \Delta v_\alpha) = s_0 + \max\left(0, v_\alpha T + \frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}} \right)
+  \begin{align}
+  \dot{v}_\alpha &= a \left[1 - \left(\frac{v_\alpha}{v_0}\right)^{\delta} - \left(\frac{s^*(v_\alpha,\Delta v_\alpha)}{s_\alpha}\right)^{2}\right]\\
+  s_\alpha &= x_{\alpha-1}-x_\alpha=x_{l}-x_\alpha\\
+  \Delta v_\alpha &=v_\alpha-v_{\alpha-1}=v_\alpha-v_l\\
+  s^*(v_\alpha, \Delta v_\alpha) &= s_0 + \max\left(0, v_\alpha T + \frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}} \right)
+  \end{align}
 \end{mymathbox}
+\paragraph{}
+The first term in Eq (1) is the \textit{free acceleration}\---the acceleration that a vehicle would take if on an open road, without considerations of desired speed or gap.
+The second term in Eq (1) is the effect on the acceleration of the vehicle approaching its desired speed.
+The third term in Eq(1) is the effect on the acceleration of relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance.
+The desired gap distance given by Eq (4) the sum of two terms the steady state desired safe distance $s_0+v_\alpha T$ and a dynamical term $\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$ representing the intelligent vehicle's braking strategy.
 
 \subsubsection{Improved Intelligent Driver Model (IIDM)}
-
+\paragraph{}
+The Improved Intelligent Driver Model is an IDM model with an improved acceleration function that aims to correct three unrealistic qualities of the IDM:
+\begin{enumerate}
+  \item 
+  If a vehicle exceeds the desired speed, the IDM acceleration function will return a very large negative acceleration.
+This is unrealistic since drivers on a highway tend to oscillate around their desired speed.
+  \item
+  If a vehicle is close to the desired speed, the desired gap $s^*(v_\alpha,\Delta v_\alpha)$ becomes much greater than $s_0 + v_\alpha T$.
+This causes the gaps between vehicles to be unrealistically large and can preclude vehicles from reaching the desired velocities.
+  \item
+  If the actual gap is less than the desired gap distance (e.g.
+when another vehicle merges into the particular vehicle's lane), then the acceleration function will return an excessively large negative acceleration.
+This unrealistic as drivers typically brake lightly to increase the gap.
+\end{enumerate}
+\begin{mymathbox}[ams gather, title=IIDM Governing Functions, colframe=blue!30!black]
+  \begin{align}
+  \frac{dv_\alpha}{dt}\Bigr|_{v_\alpha\le v_0}&= 
+  \begin{cases}
+    a (1-z^2) & z_\alpha \ge 1\\
+   a_{\text{free},\alpha}(1 - z_\alpha^{(2a)/a_{\text{free},\alpha}})& \text{otherwise}
+  \end{cases}
+  \\
+  \frac{dv_\alpha}{dt}\Bigr|_{v_\alpha> v_0}&= 
+  \begin{cases}
+    a_{\text{free},\alpha} + a (1-z_\alpha^2) & z_\alpha \ge 1\\
+    a_{\text{free},\alpha} & \text{otherwise}
+  \end{cases}\\
+  a_{\text{free},\alpha}(v_\alpha)&= \begin{cases}
+  a \left[ 1 - (\frac{v_\alpha}{v_0})^\delta \right] & v_\alpha \le v_0\\
+  -b \left[ 1 - (\frac{v_0}{v_\alpha})^{a\delta/b} \right] & v_\alpha > v_0
+  \end{cases}\\
+  z_\alpha&= \frac{s^*(v_\alpha, \Delta v_\alpha)}{s_\alpha}
+  \end{align}
+\end{mymathbox}
+\paragraph{}
+The IIDM governing equations seek to improve the unrealistic qualities of the IDM model.
+Eq (8) is the free acceleration equation, which limits the deceleration's magnitude when $v_\alpha>v_0$ to be at most b.
+To improve the acceleration function near the desired speed, we use $z_\alpha$ to indicate whether the desired gap distance is greater than the current gap distance or not.
+When $v_\alpha < v_0$, the IIDM's acceleration function (Eq (5) and Eq (6)) ensures that the actual gap is no larger than the minimum safe gap $s_\alpha = s_0+vT$.
+To maintain the intelligent and comfortable brake method of the IDM, the improved acceleration function only changes when the actual gap is near the desired gap or when the actual velocity is greater than the desired velocity.
 \subsubsection{Adaptive Cruise Control Model (ACC)}
@@
 \paragraph{}
-$$
-a_{free}(v)= \begin{cases}
-	a \left[ 1 - (\frac{v}{v_0})^\delta \right] & v \le v_0\\
-	-b \left[ 1 - (\frac{v_0}{v})^{a\delta/b} \right] & v > v_0
-\end{cases}
-$$
-
-$$\frac{dv}{dt}\Bigr|_{v\le v_0}= \begin{cases}
-a (1-z^2) & z = \frac{s*(v, \Delta v}{s} \ge 1\\
-a_{free}(1 - z^{(2a)/a_{free}})& otherwise
-\end{cases}$$
-
-$$\frac{dv}{dt}\Bigr|_{v> v_0}= \begin{cases}
-a_{free} + a (1-z^2) & z(v, \Delta v) \ge 1\\
-a_{free} & otherwise
-\end{cases}$$
-
 $$a_{CAH}(s,v,v_l, \dot{v}_l)= \begin{cases}

### 2016-12-11 project_paper/paper.tex
added results to paper

@@
 \usepackage{graphicx}
+\usepackage{float}
 
@@
 
-$$u(t-Tr) = ru_{i-j-1} + (1-r)u_{i-j},  \, \, j = int\left(\frac{T_r}{\Delta t} \right), \, \, r = \frac{T_r}{\Delta t} - j$$
+$$u(t-Tr) = ru_{i-j-1} + (1-r)u_{i-j},  \, \, j = \textrm{int}\left(\frac{T_r}{\Delta t} \right), \, \, r = \frac{T_r}{\Delta t} - j$$
 
-$$ln s^{est} - ln s = V_s w_s(t) $$
+$$\textrm{ln} s^{est} - \textrm{ln} s = V_s w_s(t) $$
 
@@
 
-% \begin{figure}[H]
-% \centering
-% \includegraphics[keepaspectratio=true,scale=0.6]{problem3d}
-% \caption{Plot of temperature cross section in pipe at various time-steps of simulation}\label{visina8}
-% \end{figure}
-
-% \begin{figure}[H]
-% \centering
-% \includegraphics[keepaspectratio=true,scale=0.6]{problem3d}
-% \caption{Plot of temperature cross section in pipe at various time-steps of simulation}\label{visina8}
-% \end{figure}
-
 \paragraph{}We then moved on to simulations of a platoon of cars with some cars being controlled by IDM-type models (in particular, Adaptive Cruise Control) and some cars being controlled by multi-car lookahead HDM models.
@@
 We find that in general, as one would expect, a larger proportion of IDM-controlled vehicles results in a quicker dissipation of the perturbation as well as a lowered probability of a vehicle collision.\\
-\textbf{RESULTS HERE}
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{maximise_distance.png}}
+  \caption{System of IDM cars on a circular track optimizing for maximum distance traveled. The more aggressive driving of the cars results in wave formation.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{optimise_comfort.png}}
+  \caption{System of IDM cars on a circular track optimizing for maximum 'comfort' (minimum jerk). As one might expect, they choose to go rather slowly and thereby achieve a very stable velocity.}
+\end{figure}
 
@@
 However, for small time steps, the `error` for 4th and 5th order methods converges, as the \texttt{odeint} solver presumably decides that 4th order is sufficiently accurate at these time steps and stops using higher order methods.\\
-\textbf{RESULTS HERE}
+\begin{figure}
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{error.png}}
+  \caption{Error to odeint decreases with accuracy of finite differencing methods.}
+\end{figure}
 \paragraph{}For our purposes, we desired to use higher order methods such that we could use larger timesteps and hopefully decrease running time of our programs while still maintaining stability in the differential equation systems.
-To display the effectiveness of these higher order methods, we ran one of our simple models for a variety of time steps over the same total time with the different solvers and noted the number of time steps at which each method stabilized.\\
-\textbf{RESULTS HERE}
+To display the effectiveness of these higher order methods, we ran one of our simple models for a variety of time steps over the same total time with the different solvers and noted the number of time steps at which each method stabilized.
+We notice here that we don't see an additional benefit from using RK5, which in this case we suspect is because the extremely complicated coefficients are contributing to existing numerical overflow issues.\\
+\begin{table}[H]
+\centering
+\caption{Timesteps for Stability with T=500s}
+\label{my-label}
+\begin{tabular}{lc}
+ Method &  Minimum Timesteps for Stability\\ \hline
+ Heun's Method & 325 \\
+ Runge-Kutta 3 &  275\\
+ Runge-Kutta 4 &  250\\
+ Runge-Kutta 5 & 250
+\end{tabular}
+\end{table}
 
@@
 
-%as a function of the reaction time Tr and the attention span ?t depends strongly on
-%the number of considered vehicles. When considering na = 5 leaders, the critical
-%effective reaction time Tr + ?t/2 at the stability limit is about twice as large as
-%the corresponding critical value without multi-anticipation (na = 1). Particularly,
-%traffic can be stable even if the reaction time exceeds the average time headway. This
-%agrees with everyday observations but cannot be realized in simulations without
-%multi-anticipation. There are limits, however: Anticipating more than five vehicles
-%ahead will change the dynamics insignificantly
+The concept of string stability refers to whether a perturbation to the beginning of the system will result in a wave that dissipates or is enhanced as it propagates through the system.
+As we have seen in previous experiments that the IDM models are excellent for adjusting to and dissipating traffic perturbations, we expect them to be string stable and indeed they are.
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{displacement_and_velocity_plot_model0_ss.png}}
+  \caption{Plot of cars 1,2,3,6,10, and 20 for IDM. String stable.}
+\end{figure}
+
+For HDM models, the presence or lack thereof of string stability depends on a number of parameters: number of leading cars being considered, time step ($\Delta t$), and reaction time ($T_r$).
+In general, string stability increases with number of leading cars being considered, although the marginal contribution of each car decreases to the point where there is not much additional benefit beyond 5 cars.
+String stability is inversely related to time step and reaction time, although reaction time has a greater effect.
+Below, we show two plots for parameters where the HDM system is string stable and string unstable.
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{HDM_anticipate_2_ss_1000_unstable.png}}
+  \caption{Plot of cars 1,2,3,6,10, and 20 for HDM with two car lookahead.
+We see that for parameters $T_r$ = .6 and $\Delta t$ = .2, the system is string unstable, as evidenced by the large wave in car 20's velocity.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{HDM_anticipate_13002_ss_stable.png}}
+  \caption{Plot of cars 1,2,3,6,10, and 20 with two car lookahead. We see that for parameters $T_r$ = .2 and $\Delta t$ = .15, the system is string stable.}
+\end{figure}
 
@@
 \paragraph{}We then attempted to optimize behavior for a single driver and were interested to see what implications this would have for the system as a whole.
-However, unsurprisingly, the following behaviors in the IDM are such that a single driver can tailgate and maximize target speed basically as much as we will allow without ever disrupting the rest of the system.
+However, the following behaviors in the IDM are such that a single driver can tailgate and maximize target speed basically as much as we will allow without ever disrupting the rest of the system.
 Even when adding some stochasticity to the lead car, we were unable to see the rest of the system suffer from traffic waves due to the behavior of the second car, because that behavior would simply be quickly dissipated by the third car.
 In fact, it seems to be true generally of these systems that unless the aggressive driving behavior of a car causes other cars to collide behind it, the greater distance achieved by one car inevitably results in other cars also achieving a greater distance/average velocity.
-However, this is likely to occur with much greater volatility in the velocity of the cars, i.e.
-a less pleasant ride.
-Then, to quantify how the erratic driving of a single car may result in penalties to the system as a whole but not to that car, we must analyze the cost to the other cars in terms of the volatility of their ride speed and possible collisions.\\
-\textbf{RESULTS HERE}
+\\
+
 \paragraph{}Next, we attempted to replicate these experiments with the more complicated Human Driver Model.
@@
 \\
-\textbf{RESULTS HERE}
+Again, we find that in general maximizing displacement for a single car leads to greater displacement for the rest of the system unless the driving behavior is so aggressive that it results in a crash.
+Given that we have seen that the Human Driver Model is more likely to propagate perturbations, we had expected that there might be in this model a penalty to other cars in greater volatility of velocity, i.e.
+a less pleasant ride, but this effect was also not observed in the data.
+We conclude that the model is largely insensitive to individual parameters in this context and that the structure of the model tends not to penalize aggressive driving unless a collision results.\\
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{optimize_s0_total.png}}
+  \caption{Changing Safe Following Distance: Changes are minimal but overall distance increases for both the car in question and the system of cars with smaller distance.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{optimize_s0.png}}
+  \caption{Variation in Distance Traveled when Changing Safe Following Distance From Initial Value of 3.0}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{optimize_t.png}}
+  \caption{Changing Safe Following Time: Changes are minimal but overall distance increases for both the car in question and the system of cars with smaller.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{optimize_t_change.png}}
+  \caption{Variation in Distance Traveled when Changing Safe Following Time From Initial Value of 1.0}
+\end{figure}
+
 \section{TODO}
+fuck this to do list, we've done enough. \\
 develop a cost function for emissions/energy usage and optimize over that \\
 can we find the parameters that will cause a phantom traffic jam?
-
-\section{Traffic Flow Simulation: Microscopic models}
-
-\section{Models and Methods}
-
-\section{Results}
 

### 2016-12-11 project_paper/paper.tex
syncing up over lunch

@@
 % reference calls and reference-list numbers.
-
-\usepackage{scicite}
 
@@
 
-\section{Macroscopic Perspective of Single Lane Traffic Interactions}
-\paragraph{}There are generally two ways to model traffic.
+\paragraph{}There are generally two ways to model traffic: from a microscopic and from a macroscopic perspective \cite[Chapter~34]{microscopic_modelling}.
 In the macroscopic perspective, traffic is viewed as a fluid or gas with a given maximum density moving according to the laws of conservation of mass.
@@
 Time-continuous microscopic models, commonly referred to as car-following models, model the characteristics of individual vehicles such as acceleration, velocity, position using a systems of coupled ordinary differential equations (ODEs).
-Car-following models are constained by assumptions that mimic realistic driving behavior such as a vehicle wants to maintain a safe driving gap between itself and vehicle in front of it (commonly called the \textit{leading vehicle}), a vehicle has a certain desired velocity, or there is a maximum comfortable acceleration.
+Car-following models are constained by assumptions that mimic realistic driving behavior such as a vehicle wants to maintain a safe driving gap between itself and vehicle in front of it (commonly called the \textit{leading vehicle}), a vehicle has a certain desired velocity, or there is a maximum comfortable acceleration \cite[chapter~10]{treiber_kesting_2013}.
 In the section, we consider a few variations of the Intelligent Driver Model (IDM), as well as Human Driver Model (HDM).
-We chose the IDM over other car-following models such as Gipps Model because the IDM has direct extensions to modeling human driving behavior using the HDM.
+%We chose the IDM over other car-following models such as Gipps Model because the IDM has direct extensions to modeling human driving behavior using the HDM.
+We chose the IDM over other car-following models such as the Gipps Model or Newell's due to the oversimplification of some acceleration assumptions that the alternative models make\cite[chapter~11]{treiber_kesting_2013}.
+The result is that the IDM provides the simplest implementation of plausible driving characteristics.
 \subsection{Intelligent Driver Models}
@@
 The third term in Eq(1) is the effect on the acceleration of relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance.
-The desired gap distance given by Eq (4) the sum of two terms the steady state desired safe distance $s_0+v_\alpha T$ and a dynamical term $\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$ representing the intelligent vehicle's braking strategy.
+The desired gap distance given by Eq (4) the sum of two terms the steady state desired safe distance $s_0+v_\alpha T$ and a dynamical term $\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$ representing the intelligent vehicle's braking strategy \cite{intelligent_driver_model}.
 
@@
 % style.
+\newpage
 
-\bibliography{scibib}
-
-\bibliographystyle{Science}
+\bibliography{bibliography}
+\bibliographystyle{ieeetr}
 

### 2016-12-11 project_paper/paper.tex
Added ACC, HDM sections

@@
 \paragraph{}
-The Improved Intelligent Driver Model is an IDM model with an improved acceleration function that aims to correct three unrealistic qualities of the IDM:
+The Improved Intelligent Driver Model is an IDM model with an improved acceleration function that aims to correct two unrealistic qualities of the IDM when a vehicle is near the desired speed:
 \begin{enumerate}
@@
 This causes the gaps between vehicles to be unrealistically large and can preclude vehicles from reaching the desired velocities.
-  \item
-  If the actual gap is less than the desired gap distance (e.g.
-when another vehicle merges into the particular vehicle's lane), then the acceleration function will return an excessively large negative acceleration.
-This unrealistic as drivers typically brake lightly to increase the gap.
 \end{enumerate}
@@
   \begin{cases}
-    a (1-z^2) & z_\alpha \ge 1\\
+    a (1-z_\alpha^2) & z_\alpha \ge 1\\
    a_{\text{free},\alpha}(1 - z_\alpha^{(2a)/a_{\text{free},\alpha}})& \text{otherwise}
@@
 \subsubsection{Adaptive Cruise Control Model (ACC)}
+\paragraph{}
+The ACC is an improvement on the IIDM model that is designed so that a vehicle can handle is discontinuous changes in the gap between itself and the leading vehicle or discontinuous changes in the speed of the leading.
+One scenario where discontinuous changes in speed or gap would occur is when the leading vehicle either enters or exits the lane.
+Since the IDM based models are only meant for controlling the acceleration of a vehicle within a single lane, this scenario is equivalent to the leading vehicle simply vanishing into or appearing out of thin air.
+In the IDM and IIDM models, vehicles overreact if the gap is suddenly decreased because the models are designed to be accident free under all circumstances.
+However, a vehicle slamming on the breaks when the gap is decreased is not realistic and can even be dangerous since most human drivers do not anticipate sudden changes in the acceleration of the vehicle in front of it.
+It is important note that the ACC model is not guaranteed to be accident-free; the driving style is less defensive by by design.
+\paragraph{}
+To attenuate the semi-autonomous vehicle's reaction to sudden decreases in the gap distance, vehicles in the ACC model act according to a constant-acceleration heuristic, rather than making decisions based off the worst-case scenario.
+The constant-acceleration heuristic (CAH) assumes that vehicles will have constant acceleration in the near future, no safe distance or time gap is necessary at any time, and vehicles will react instantaneously.
+\begin{mymathbox}[ams gather, title=ACC Governing Functions, colframe=blue!30!black]
+  \begin{align}
+  a_{\text{ACC},\alpha}&= 
+  \begin{cases}
+  a_{\text{IIDM},\alpha} & a_{\text{IIDM},\alpha} \ge a_{\text{CAH},\alpha}\\
+  (1-c)a_{\text{IIDM},\alpha} + c\left[a_{\text{CAH},\alpha} + b \tanh(\frac{a_{\text{IIDM},\alpha}-a_{\text{CAH},\alpha}}{b}) \right] & \text{otherwise}
+  \end{cases}\\
+  a_{\text{CAH},\alpha}&= 
+  \begin{cases}
+  \frac{v_\alpha^2\tilde{a}_l}{v_l^2 - 2 s_\alpha\tilde{a}_l} & v_l(v_\alpha-v_l) \le -2s_\alpha\tilde{a}_l\\
+  \tilde{a}_l - \frac{(v_\alpha-v_l)^2 \Theta (v_\alpha-v_l)}{2s_\alpha} & \text{otherwise}
+  \end{cases}\\
+  \tilde{a}_l &= \min(\dot{v}_l, a)\\
+   \Theta (x)&= 
+   \begin{cases}
+   1 & x\geq 0\\
+   0 & \text{otherwise}
+   \end{cases}
+   \end{align}
+\end{mymathbox}
+\paragraph{}
+Under the constant-acceleration heuristic, the vehicle's acceleration is given by Eq (9), where $\tilde{a}_l$ is the effective acceleration of the leading vehicle (limited by the maximum acceleration parameter).
+One important distinction between the ACC and the previous intelligent models is that the ACC uses the leading vehicle's acceleration as an input, in addition to the leading vehicle's speed and gap.
+However, the ACC model only uses the CAH acceleration as an indicator of whether the IIDM acceleration is overreacting to the situation.
+Hence, the ACC model also calculates the acceleration function from the IIDM model.
+The acceleration function for the ACC model is given by Eq (9), where $c$ is the "coolness parameter", which is used to tune the ACC models behavior: $c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model.
+If the IIDM acceleration is greater than or equal to the CAH acceleration, then the ACC model uses the IIDM acceleration, since the IIDM model is accident-free and thus safe.
+Otherwise, the AAC model uses a less cautious acceleration function.
+\section{Microscopic Perspective of Traffic Interaction: Human Driver Models}
+\paragraph{}
+The Human Driver Model (HDM) is an extension of the IDM model that models human driving behavior by making five important changes to the IDM:
+\begin{enumerate}
+  \item The\textit{ reaction times} are no longer zero.
+Humans have delayed reactions.
+The HDM requires a reaction time parameter $T_r$, which is the delay in response to stimulus.
+Hence, the system of ODEs for the HDM is a system of time-delayed differential equations.
+  \item For a given driver, the exact distance to and speed of the leading vehicle are not known.
+The driver estimates these quantities with inherent \textit{estimation error}.
+Note that the HDM assumes that the driver knows their own velocity, since the vehicle has a speedometer.
+  \item Drivers do not update their acceleration exactly to the value given by the acceleration function.
+This is because a driver controls the car by operating the gas and brake pedals, which are imprecise and induce \textit{driver control errors}.
+  \item Experienced drivers \textit{anticipate} what will happen in the near future and use these anticipations to inform their own behavior.
+  \item Human drivers are able to look ahead multiple vehicles in a single lane of traffic.
+For example, a human driver often can see brake lights of a vehicle 3 cars ahead on a highway and will react to those brake lights in the distance.
+\textit{multiple anticipation} allowing the human drivers update their acceleration using information from multiple cars ahead.
+\end{enumerate}
+\begin{mymathbox}[title=HDM Parameters,colframe=blue!30!black]
+  $T_r$: human reaction time\\
+  $\tilde{\tau}$: persistence time for Wiener process autocorrelation function\\
+  $V_s$: variation coefficient\---standard deviation of the relative difference between $s_\alpha^\text{est}$ and $s_\alpha$\\
+  $\sigma_r$: standard deviation of the relative approach rate\\
+  $\sigma_a$: standard deviation of the acceleration error
+\end{mymathbox}
 
-\paragraph{}
-$$a_{CAH}(s,v,v_l, \dot{v}_l)= \begin{cases}
-\frac{v^2\tilde{a}_l}{v_l^2 - 2 \tilde(a)_l} & v_l(v-v_l) \le -2s\tilde{a}_l\\
-\tilde{a}_l - \frac{(v-v_l)^2 \Theta (v-v_l)}{2s} & otherwise
-\end{cases}$$
-
-$$\tilde{a}_l(\dot(v)_l) = min(\dot{v}_l, a)$$
-
-$$a_{ACC}= \begin{cases}
-a_{IIDM} & a_{IIDM} \ge a_{CAH}\\
-(1-c)a_{IIDM} + c\left[a_{CAH} + b tanh(\frac{a_{IIDM}-a_{CAH}}{b}) \right] & otherwise
-\end{cases}$$
-
-$$\dot{v}(t) = a_{mic}\left[s(t-T_r), v(t-T_r), v_l(t-T_r) \right]$$
-
-$$u(t-Tr) = ru_{i-j-1} + (1-r)u_{i-j},  \, \, j = \textrm{int}\left(\frac{T_r}{\Delta t} \right), \, \, r = \frac{T_r}{\Delta t} - j$$
-
-$$\textrm{ln} s^{est} - \textrm{ln} s = V_s w_s(t) $$
-
-$$v^{est} - v_l = -s \sigma_r w_l(t)$$
-
-%This numerical analysis can be contrasted to the analytical solution that was presented in (**another section 2?**).\par
-%A car`s speed is governed by the laws of motion:
-%\begin{equation}
-%v_f = v_i + a\Delta t
-%\end{equation}
-%
-%An individual car can be modeled as such over a 100,000m track and asked to travel at $v_{target} = 60miles.h^{-1}$ (i.e.
-$26.8m.s^{-1}$).
-As the driver is not an perfect controller he is unable to hold the speed exactly at $v_{target}$.
-We can therefore model his fluctuations as random acceleration with the corresponding adjustments to correct the acceleration:
-%\begin{equation}
-%a_{correction_i} = \alpha(v_{instant} - v_{target}) + N(0,\sigma_1)
-%\end{equation}
-%
-%We note that for more than one car in a stream of traffic the car's acceleration is then also dependent on the car ahead of it:
-%\begin{equation}\label{eq:interaction}
-%a_{interaction_i} = \beta (s_{instant} - s_{target})
-%\end{equation}
-%
-%In equation \ref{eq:interaction}, we see that the acceleration of the car is also dependent on the distance between any one car and the car that is directly ahead of it.
-The final acceleration of the car is therefore:
-%\begin{equation}\label{eq:acceleration}
-%a_{i} = a_{correction_i} + a_{interaction_i}
-%\end{equation}
-%
-%If we assume the cars should travel at $v_{target}$ and that they should maintain an $s_{target} = 5m$ spacing between them, we can then model the velocity of the car as:
-%
-%\begin{equation}\label{eq:velocity}
-%v_{i} = \frac{d(a_i)}{dt}
-%\end{equation}
-
-\section{Microscopic Perspective of Traffic Interaction : Human Driver Models}
-
+%\begin{mymathbox}[ams gather, title=HDM Governing Equations,colframe=blue!30!black]
+%  \begin{align}
+%\dot{v}_\alpha(t) &= a_{\alpha}\left[s_\alpha(t-T_r), v_\alpha(t-T_r), v_l(t-T_r) \right]\\
+%u(t-T_r) &= ru_{i-j-1} + (1-r)u_{i-j},  \, \, j = \textrm{int}\left(\frac{T_r}{\Delta t} \right), \, \, r = \frac{T_r}{\Delta t} - j\\
+%\langle w(t)w(t')\rangle&  =\exp\bigg(-\frac{|t-t'|}{\tilde{\tau}}\bigg)\\
+%w_i&=e^{\frac{-\Delta t}{\tilde{\tau}}}w_{i-1}+\sqrt{\frac{2\Delta t}{\tilde\tau}}\eta_i\\
+%\ln s_\alpha^{\text{est}} - \ln s_\alpha &= V_s w_{s,\alpha}(t)\\
+%v^{est} - v_l &= -s_\alpha \sigma_r w_{l,\alpha}(t)\\
+%\dot v_\alpha(t)&=a_\alpha(s_\alpha^\text{est},v_\alpha,v_l^\text{est})+\sigma_a w_{a,\alpha}(t)\\
+%v_\alpha^{\text{prog}}(t)&=v_\alpha(t-T_r)+T_r\dot v_\alpha(t-T_r)\\
+%v_l^{\text{prog}}(t)&=v_l^\text{est}(t-T_r)\\
+%s_\alpha^{\text{prog}}(t)&=s_\alpha^\text{est}(t-T_r)-T_r\Delta v_\alpha^\text{est}(t-T_r)
+%  \end{align}
+%\end{mymathbox}
+\paragraph{1. Delayed Reaction Times:}
+The HDM incorporates human reaction time by introducing a new reaction time parameter $T_r$.
+Vehicles in the model then update their acceleration based delayed information (i.e.
+the state of the system $T_r$ seconds before) as shown in Eq (13).
+Hence, the HDM is based off time-delayed differential equations.
+Eq (13) cannot be solved analytically, but $u(t-T_r)$ can be approximated numerically by linear interpolation using the scheme in Eq (14), where $u$ can be any of $s_\alpha,v_\alpha,\Delta v_\alpha,v_l$, $\Delta t$ is the update time of the scheme, and $i=\frac{t}{\Delta t}$ is the update step number.
+\begin{mymathbox}[ams gather, title=Delayed Reaction Time Equations Equations,colframe=blue!30!black]
+  \begin{align}
+  \dot{v}_\alpha(t) &= a_{\alpha}\left[s_\alpha(t-T_r), v_\alpha(t-T_r), v_l(t-T_r) \right]\\
+  u(t-T_r) &= ru_{i-j-1} + (1-r)u_{i-j},  \, \, j = \textrm{int}\left(\frac{T_r}{\Delta t} \right), \, \, r = \frac{T_r}{\Delta t} - j
+  \end{align}
+\end{mymathbox}
+\paragraph{2. Estimation Errors:}
+The HDM addresses a driver's error in estimating the distance to and speed of the vehicle in front of it by adding Gaussian noise to the true value of the estimand.
+However, it is likely a particular driver's estimation error will be correlated over time.
+Hence, for each driver $\alpha$, the HDM two stochastic processes ($w_{s,\alpha}(t),w_{l,\alpha}(t)$) to represent the driver's estimation errors of the distance to and speed of the leading vehicle at each timestep, respectively.
+Each of the stochastic processes is independent and comes from the standard Gaussian distribution.
+Moreover, each stochastic process uses the autocorrelation function given by Eq (15), which models how a driver's estimation errors are likely correlated with other estimation errors at other nearby times.
+\textbf{SHOULD I INCLUDE STOCHASTIC DIFF EQ FOR WIENER PROCESS?} Each stochastic process is created approximating a Wiener Process using the numerical scheme given in Eq (16), where $i=\frac{t}{\Delta t}$ is the update step number and $\eta_i \sim N(0,1)$.
+Using these stochastic processes, the estimation errors (or log errors) of distance to and speed of the leading vehicle are given by Eq (17, 18), respectively, where $s_\alpha^\text{est}$ and $v_l^\text{est}$ are the estimates.
+\begin{mymathbox}[ams gather, title=Estimation Error Equations,colframe=blue!30!black]
+  \begin{align}
+  \langle w(t)w(t')\rangle&  =\exp\bigg(-\frac{|t-t'|}{\tilde{\tau}}\bigg)\\
+  w_i&=e^{\frac{-\Delta t}{\tilde{\tau}}}w_{i-1}+\sqrt{\frac{2\Delta t}{\tilde\tau}}\eta_i\\
+  \ln s_\alpha^{\text{est}} - \ln s_\alpha &= V_s w_{s,\alpha}(t)\\
+  v^{est} - v_l &= -s_\alpha \sigma_r w_{l,\alpha}(t)
+  \end{align}
+\end{mymathbox}
+\paragraph{3. Driver Control Errors:}
+The HDM handles driver control errors adding noise to the acceleration function at each time step.
+Specifically, another stochastic process $w_{a,\alpha}(t)$ with the same properties as $w_{s,\alpha}(t)$ and  $w_{l,\alpha}(t)$ is created for each driver.
+This noise is added to the acceleration function as shown in the Eq (19).
+\begin{mymathbox}[ams gather, title=Driver Control Error Equation Equations,colframe=blue!30!black]
+  \begin{align}
+\dot v_\alpha(t)&=a_\alpha(s_\alpha^\text{est},v_\alpha,v_l^\text{est})+\sigma_a w_{a,\alpha}(t)
+  \end{align}
+\end{mymathbox}
+\paragraph{4. Temporal Anticipation:}
+The HDM models an experienced driver's anticipation by having driver's assume that their acceleration is constant over the reaction time $T_r$.
+This assumption allows a driver to predict what their own speed will be $T_r$ seconds in the future.
+This prediction is given by Eq (20).
+Moreover, in the HDM, a driver assumes that the velocity of leading vehicle is constant over next $T_r$ seconds.
+This allows the driver to make predictions of the distance to and speed of the leading vehicle, as shown in Eq (21, 22) respectively, where $\Delta v_\alpha^\text{est}=v_\alpha-v_l^\text{est}$.
+\begin{mymathbox}[ams gather, title=Temporal Anticipation Equations,colframe=blue!30!black]
+  \begin{align}
+  v_\alpha^{\text{prog}}(t)&=v_\alpha(t-T_r)+T_r\dot v_\alpha(t-T_r)\\
+  v_l^{\text{prog}}(t)&=v_l^\text{est}(t-T_r)\\
+  s_\alpha^{\text{prog}}(t)&=s_\alpha^\text{est}(t-T_r)-T_r\Delta v_\alpha^\text{est}(t-T_r)
+  \end{align}
+\end{mymathbox}
+\paragraph{5. Multi-Vehicle Anticipation:}
+The HDM models allows a driver to be aware multiple vehicles ahead.
+The acceleration function (shown in Eq(23)) is broken up into two terms: a free acceleration term that representing the acceleration on an open road, and an interaction term that models the effect of the driver's awareness of multiple vehicles in front of it on the driver's acceleration.
+The free acceleration and interaction terms directly follow from adapting the IDM acceleration function to include the distances to and speeds of multiple leading cars.
+The coefficient of the interaction term, $c_{\text{IDM}}$, is used to keep the interaction term from blowing up in the negative direction as $n_a$ increases.
+\begin{mymathbox}[ams gather, title=Multi-Vehicle Anticipation Equations,colframe=blue!30!black]
+  \begin{align}
+  \dot v_\alpha &= a_{\text{free},\alpha}+c_{\text{IDM}}\sum_{\beta=\alpha-n_a}^{\alpha-1}a_{\text{int}}(s_{\alpha\beta},v_\alpha,v_\beta)\\
+  a^{\text{IDM}}_{\text{free},\alpha}&=a\Bigg[1-\bigg(\frac{v}{v_0}\bigg)^\delta\Bigg]\\
+  a^{\text{IDM}}_{\text{int}}(s_{\alpha\beta},v_\alpha,v_\beta)&=-a\Bigg(\frac{s^*(v_\alpha,v_\alpha-v_\beta)}{s_{\alpha\beta}}\Bigg)^2\\
+  s_{\alpha\beta}&=\sum_{j=0}^{\alpha-\beta-1}s_{\alpha-j}\\
+  c_{\text{IDM}}&=\Bigg(\sum_{j=1}^{n_a}\frac{1}{j^2}\Bigg)^{-1}
+  \end{align}
+\end{mymathbox}
+\paragraph{HDM Model:}
+Assembling the various components of the HDM model, the HDM governing equation is obtained (Eq (28)).
+\begin{mymathbox}[ams gather, title=HDM Governing Equation,colframe=blue!30!black]
+  \begin{align}
+\dot v_\alpha &= a_{\text{free},\alpha}^{\text{IDM}}+c_{\text{IDM}}\sum_{\beta=\alpha-n_a}^{\alpha-1}a_{\text{int}}^{\text{IDM}}(s^{\text{prog}}_{\alpha\beta},v^{\text{prog}}_\alpha,v^{\text{prog}}_\beta)
+  \end{align}
+\end{mymathbox}
 \section{Numerical Simulations of Traffic Systems}

### 2016-12-11 project_paper/paper.tex
updated images

@@
 
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{displacement_and_velocity_plot_model0.pdf}}
+  \caption{System of 50 IDM-controlled cars on a circular track.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{displacement_and_velocity_plot_model2.pdf}}
+  \caption{System of 50 ACC-controlled cars on a circular track.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{HDM_anticipate_1.pdf}}
+  \caption{System of 50 HDM-controlled cars with one car-lookahead on a circular track.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{HDM_anticipate_5.pdf}}
+  \caption{System of 50 HDM-controlled cars with five car-lookahead on a circular track. Note the significant decrease in volatility of velocity.}
+\end{figure}
+
 \paragraph{}We then moved on to simulations of a platoon of cars with some cars being controlled by IDM-type models (in particular, Adaptive Cruise Control) and some cars being controlled by multi-car lookahead HDM models.
@@
 We find that in general, as one would expect, a larger proportion of IDM-controlled vehicles results in a quicker dissipation of the perturbation as well as a lowered probability of a vehicle collision.\\
+
 \begin{figure}[H]
   \centering
-  \fbox{\includegraphics[width=0.95\textwidth]{maximise_distance.png}}
-  \caption{System of IDM cars on a circular track optimizing for maximum distance traveled. The more aggressive driving of the cars results in wave formation.}
+  \fbox{\includegraphics[width=0.95\textwidth]{blended_3_35_127pm.png}}
+  \caption{System with 35 \% IDM cars on a straight road with acceleration of first car perturbed. System maintains a high degree of volatility in velocity long after the initial perturbation.}
 \end{figure}
@@
   \centering
-  \fbox{\includegraphics[width=0.95\textwidth]{optimise_comfort.png}}
-  \caption{System of IDM cars on a circular track optimizing for maximum 'comfort' (minimum jerk). As one might expect, they choose to go rather slowly and thereby achieve a very stable velocity.}
+  \fbox{\includegraphics[width=0.95\textwidth]{blended_3_90_127pm.png}}
+  \caption{System with 90 \% IDM cars on a straight road with acceleration of first car perturbed. System quickly dissipates the majority of the volatility in velocity.}
 \end{figure}
@@
 However, for small time steps, the `error` for 4th and 5th order methods converges, as the \texttt{odeint} solver presumably decides that 4th order is sufficiently accurate at these time steps and stops using higher order methods.\\
-\begin{figure}
+\begin{figure}[H]
   \centering
@@
 \\
-\textbf{RESULTS HERE}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{maximise_distance.png}}
+  \caption{System of IDM cars on a circular track optimizing for maximum distance traveled. The more aggressive driving of the cars results in wave formation.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \fbox{\includegraphics[width=0.95\textwidth]{optimise_comfort.png}}
+  \caption{System of IDM cars on a circular track optimizing for maximum 'comfort' (minimum jerk). As one might expect, they choose to go rather slowly and thereby achieve a very stable velocity.}
+\end{figure}
+
 \paragraph{}We then attempted to optimize behavior for a single driver and were interested to see what implications this would have for the system as a whole.
@@
   \fbox{\includegraphics[width=0.95\textwidth]{optimize_s0.png}}
-  \caption{Variation in Distance Traveled when Changing Safe Following Distance From Initial Value of 3.0}
+  \caption{Variation in Distance Traveled when Changing Safe Following Distance Relative to Initial Value of 3.0}
 \end{figure}
@@
   \fbox{\includegraphics[width=0.95\textwidth]{optimize_t_change.png}}
-  \caption{Variation in Distance Traveled when Changing Safe Following Time From Initial Value of 1.0}
+  \caption{Variation in Distance Traveled when Changing Safe Following Time Relative to Initial Value of 1.0}
 \end{figure}

### 2016-12-11 project_paper/paper.tex
revisions

@@
 The IDM is a simple, accident-free model that can model traffic systems under all realistic traffic conditions.
-The acceleration function, shown in Eq (1), of a particular vehicle $\alpha$ is a function the current gap between itself and the leading vehicle ($s_\alpha$), the desired speed $v_0$, and the speed of the vehicle leading vehicle $v_l$.
+The acceleration function, shown in Eq (1), of a particular vehicle $\alpha$ is a function the distance between itself and the leading vehicle ($s_\alpha$, referred to as the gap), the desired speed $v_0$, and the speed of the leading vehicle $v_l$.
 
@@
 \paragraph{}
-The first term in Eq (1) is the \textit{free acceleration}\---the acceleration that a vehicle would take if on an open road, without considerations of desired speed or gap.
-The second term in Eq (1) is the effect on the acceleration of the vehicle approaching its desired speed.
-The third term in Eq(1) is the effect on the acceleration of relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance.
-The desired gap distance given by Eq (4) the sum of two terms the steady state desired safe distance $s_0+v_\alpha T$ and a dynamical term $\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$ representing the intelligent vehicle's braking strategy \cite{intelligent_driver_model}.
+The first term in Eq (1) is the \textit{free acceleration}\---the acceleration that a vehicle would take if it were on a wide open road, without considerations of desired speed or gap.
+The second term in Eq (1) is the effect on the vehicle's acceleration as it approaches its desired speed.
+The third term in Eq(1) is the effect, on the acceleration, of relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance.
+The desired gap distance given by Eq (4) is the sum of a desired safe distance $s_0+v_\alpha T$ and a dynamical term $\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$ representing the intelligent vehicle's braking strategy \cite{intelligent_driver_model}.
 
@@
 \paragraph{}
-The Improved Intelligent Driver Model is an IDM model with an improved acceleration function that aims to correct two unrealistic qualities of the IDM when a vehicle is near the desired speed:
+The Improved Intelligent Driver Model is an IDM model with an enhanced acceleration function that aims to correct two unrealistic qualities of the IDM when a vehicle's speed is close to the desired speed:
 \begin{enumerate}
@@
   \item
-  If a vehicle is close to the desired speed, the desired gap $s^*(v_\alpha,\Delta v_\alpha)$ becomes much greater than $s_0 + v_\alpha T$.
-This causes the gaps between vehicles to be unrealistically large and can preclude vehicles from reaching the desired velocities.
+  If a vehicle is close to the desired speed, the desired gap $s^*(v_\alpha,\Delta v_\alpha)$ becomes much wider than $s_0 + v_\alpha T$.
+This is unrealistic and may preclude vehicles from reaching the desired velocity $v_0$.
 \end{enumerate}
@@
 The IIDM governing equations seek to improve the unrealistic qualities of the IDM model.
-Eq (8) is the free acceleration equation, which limits the deceleration's magnitude when $v_\alpha>v_0$ to be at most b.
-To improve the acceleration function near the desired speed, we use $z_\alpha$ to indicate whether the desired gap distance is greater than the current gap distance or not.
+Eq (8) is the free acceleration equation, which limits the deceleration's magnitude when $v_\alpha>v_0$ to be at most $b$.
+This prevents vehicles from slamming on the brakes as soon as they go faster than the desire velocity.
+To improve the acceleration function's behavior near the desired speed, the IIDM sets the acceleration differently depending on whether the speed and gap above or below their respective desired values.
 When $v_\alpha < v_0$, the IIDM's acceleration function (Eq (5) and Eq (6)) ensures that the actual gap is no larger than the minimum safe gap $s_\alpha = s_0+vT$.
-To maintain the intelligent and comfortable brake method of the IDM, the improved acceleration function only changes when the actual gap is near the desired gap or when the actual velocity is greater than the desired velocity.
+To maintain the intelligent and comfortable brake method of the IDM, this improved acceleration function mimics the behavior of the IDM acceleration function except for when the actual gap is near the desired gap or when the actual speed is greater than the desired speed.
 \subsubsection{Adaptive Cruise Control Model (ACC)}
 \paragraph{}
-The ACC is an improvement on the IIDM model that is designed so that a vehicle can handle is discontinuous changes in the gap between itself and the leading vehicle or discontinuous changes in the speed of the leading.
-One scenario where discontinuous changes in speed or gap would occur is when the leading vehicle either enters or exits the lane.
+The ACC is an improvement on the IIDM model that is designed so that a vehicle can handle discontinuous changes in the gap between itself and the leading vehicle or discontinuous changes in the speed of the leading vehicle.
+One scenario such discontinuous changes would occur is when the leading vehicle either enters or exits the lane.
 Since the IDM based models are only meant for controlling the acceleration of a vehicle within a single lane, this scenario is equivalent to the leading vehicle simply vanishing into or appearing out of thin air.
-In the IDM and IIDM models, vehicles overreact if the gap is suddenly decreased because the models are designed to be accident free under all circumstances.
-However, a vehicle slamming on the breaks when the gap is decreased is not realistic and can even be dangerous since most human drivers do not anticipate sudden changes in the acceleration of the vehicle in front of it.
-It is important note that the ACC model is not guaranteed to be accident-free; the driving style is less defensive by by design.
+In the IDM and IIDM models, vehicles overreact if the gap is suddenly decreased because the models are designed to avoid accidents under all circumstances.
+However, a vehicle slamming on the breaks when the gap is decreased is unrealistic and can even be dangerous since most human drivers do not anticipate sudden changes in the acceleration of the vehicle in front of it.
+It is important note that the ACC model is not guaranteed to be accident-free; the driving style is less defensive by design.
 \paragraph{}
 To attenuate the semi-autonomous vehicle's reaction to sudden decreases in the gap distance, vehicles in the ACC model act according to a constant-acceleration heuristic, rather than making decisions based off the worst-case scenario.
-The constant-acceleration heuristic (CAH) assumes that vehicles will have constant acceleration in the near future, no safe distance or time gap is necessary at any time, and vehicles will react instantaneously.
+The constant-acceleration heuristic (CAH) assumes that vehicles will have constant acceleration in the near future, no safe distance or time gap is necessary at any time, and that vehicles will react instantaneously.
 \begin{mymathbox}[ams gather, title=ACC Governing Functions, colframe=blue!30!black]
@@
 \paragraph{}
-Under the constant-acceleration heuristic, the vehicle's acceleration is given by Eq (9), where $\tilde{a}_l$ is the effective acceleration of the leading vehicle (limited by the maximum acceleration parameter).
-One important distinction between the ACC and the previous intelligent models is that the ACC uses the leading vehicle's acceleration as an input, in addition to the leading vehicle's speed and gap.
-However, the ACC model only uses the CAH acceleration as an indicator of whether the IIDM acceleration is overreacting to the situation.
+Under the constant-acceleration heuristic, the vehicle's acceleration is given by Eq (9), where $\tilde{a}_l$ is the effective acceleration of the leading vehicle (limited by the maximum acceleration parameter $a$).
+One important distinction between the ACC and the previous intelligent driving models is that the ACC uses the leading vehicle's acceleration as an input, in addition to the leading vehicle's speed and gap.
+It is important to note that the CAH acceleration is only used as indicator of whether the IIDM acceleration is overreacting to the situation.
 Hence, the ACC model also calculates the acceleration function from the IIDM model.
-The acceleration function for the ACC model is given by Eq (9), where $c$ is the "coolness parameter", which is used to tune the ACC models behavior: $c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model.
-If the IIDM acceleration is greater than or equal to the CAH acceleration, then the ACC model uses the IIDM acceleration, since the IIDM model is accident-free and thus safe.
+The acceleration function for the ACC model is given by Eq (9), where $c$ is the "coolness parameter", which is used to tune the ACC model's behavior: $c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model.
+If the IIDM acceleration is greater than or equal to the CAH acceleration, then the ACC model uses the IIDM acceleration, since the IIDM model is accident-free and thus guaranteed to be safe.
 Otherwise, the AAC model uses a less cautious acceleration function.
@@
 Humans have delayed reactions.
-The HDM requires a reaction time parameter $T_r$, which is the delay in response to stimulus.
-Hence, the system of ODEs for the HDM is a system of time-delayed differential equations.
+The HDM requires a reaction time parameter $T_r$, which is the delay in response to changes in the system.
+Hence, the the HDM is governed by a system of time-delayed differential equations.
   \item For a given driver, the exact distance to and speed of the leading vehicle are not known.
 The driver estimates these quantities with inherent \textit{estimation error}.
-Note that the HDM assumes that the driver knows their own velocity, since the vehicle has a speedometer.
-  \item Drivers do not update their acceleration exactly to the value given by the acceleration function.
-This is because a driver controls the car by operating the gas and brake pedals, which are imprecise and induce \textit{driver control errors}.
-  \item Experienced drivers \textit{anticipate} what will happen in the near future and use these anticipations to inform their own behavior.
+The HDM assumes that the driver knows their own velocity, since the vehicle has a speedometer.
+  \item A driver does not update their acceleration exactly to the value given by the acceleration function.
+This is because a driver controls the car by operating the gas and brake pedals, which leads to imprecise accelerations and induces \textit{driver control errors}.
+  \item Experienced drivers \textit{anticipate} what will happen in the near future and use this anticipation to inform their own behavior.
   \item Human drivers are able to look ahead multiple vehicles in a single lane of traffic.
-For example, a human driver often can see brake lights of a vehicle 3 cars ahead on a highway and will react to those brake lights in the distance.
-\textit{multiple anticipation} allowing the human drivers update their acceleration using information from multiple cars ahead.
+For example, a human driver often can see brake lights of a vehicle three cars ahead on a highway and will react to the brake lights in the distance.
+The HDM models this \textit{multi-vehicle anticipation} allowing the human drivers to update their acceleration based multiple cars ahead of them.
 \end{enumerate}
@@
 \end{mymathbox}
-
-%\begin{mymathbox}[ams gather, title=HDM Governing Equations,colframe=blue!30!black]
-%  \begin{align}
-%\dot{v}_\alpha(t) &= a_{\alpha}\left[s_\alpha(t-T_r), v_\alpha(t-T_r), v_l(t-T_r) \right]\\
-%u(t-T_r) &= ru_{i-j-1} + (1-r)u_{i-j},  \, \, j = \textrm{int}\left(\frac{T_r}{\Delta t} \right), \, \, r = \frac{T_r}{\Delta t} - j\\
-%\langle w(t)w(t')\rangle&  =\exp\bigg(-\frac{|t-t'|}{\tilde{\tau}}\bigg)\\
-%w_i&=e^{\frac{-\Delta t}{\tilde{\tau}}}w_{i-1}+\sqrt{\frac{2\Delta t}{\tilde\tau}}\eta_i\\
-%\ln s_\alpha^{\text{est}} - \ln s_\alpha &= V_s w_{s,\alpha}(t)\\
-%v^{est} - v_l &= -s_\alpha \sigma_r w_{l,\alpha}(t)\\
-%\dot v_\alpha(t)&=a_\alpha(s_\alpha^\text{est},v_\alpha,v_l^\text{est})+\sigma_a w_{a,\alpha}(t)\\
-%v_\alpha^{\text{prog}}(t)&=v_\alpha(t-T_r)+T_r\dot v_\alpha(t-T_r)\\
-%v_l^{\text{prog}}(t)&=v_l^\text{est}(t-T_r)\\
-%s_\alpha^{\text{prog}}(t)&=s_\alpha^\text{est}(t-T_r)-T_r\Delta v_\alpha^\text{est}(t-T_r)
-%  \end{align}
-%\end{mymathbox}
 \paragraph{1. Delayed Reaction Times:}
@@
 the state of the system $T_r$ seconds before) as shown in Eq (13).
-Hence, the HDM is based off time-delayed differential equations.
-Eq (13) cannot be solved analytically, but $u(t-T_r)$ can be approximated numerically by linear interpolation using the scheme in Eq (14), where $u$ can be any of $s_\alpha,v_\alpha,\Delta v_\alpha,v_l$, $\Delta t$ is the update time of the scheme, and $i=\frac{t}{\Delta t}$ is the update step number.
+Eq (13) cannot be solved analytically, but $u(t-T_r)$ can be approximated numerically by linear interpolation using the scheme in Eq (14), where $u$ is a placeholder for any of $s_\alpha,v_\alpha,\Delta v_\alpha,v_l$.
+In this numerical scheme, $\Delta t$ is the update time and $i=\frac{t}{\Delta t}$ is the update step number.
 \begin{mymathbox}[ams gather, title=Delayed Reaction Time Equations Equations,colframe=blue!30!black]
@@
 However, it is likely a particular driver's estimation error will be correlated over time.
-Hence, for each driver $\alpha$, the HDM two stochastic processes ($w_{s,\alpha}(t),w_{l,\alpha}(t)$) to represent the driver's estimation errors of the distance to and speed of the leading vehicle at each timestep, respectively.
+Hence, for each driver $\alpha$, the HDM creates two stochastic processes ($w_{s,\alpha}(t),w_{l,\alpha}(t)$) to represent the driver's error in estimating the distance to and speed of the leading vehicle at each time step, respectively.
 Each of the stochastic processes is independent and comes from the standard Gaussian distribution.
-Moreover, each stochastic process uses the autocorrelation function given by Eq (15), which models how a driver's estimation errors are likely correlated with other estimation errors at other nearby times.
-\textbf{SHOULD I INCLUDE STOCHASTIC DIFF EQ FOR WIENER PROCESS?} Each stochastic process is created approximating a Wiener Process using the numerical scheme given in Eq (16), where $i=\frac{t}{\Delta t}$ is the update step number and $\eta_i \sim N(0,1)$.
+Moreover, each stochastic process uses the autocorrelation function given by Eq (15).
+Each stochastic process is created approximating a Wiener Process using the numerical scheme given in Eq (16), where $i=\frac{t}{\Delta t}$ is the update step number and $\eta_i \sim N(0,1)$.
 Using these stochastic processes, the estimation errors (or log errors) of distance to and speed of the leading vehicle are given by Eq (17, 18), respectively, where $s_\alpha^\text{est}$ and $v_l^\text{est}$ are the estimates.
@@
 \paragraph{3. Driver Control Errors:}
-The HDM handles driver control errors adding noise to the acceleration function at each time step.
+The HDM includes driver control errors by adding noise to the acceleration function at each time step.
 Specifically, another stochastic process $w_{a,\alpha}(t)$ with the same properties as $w_{s,\alpha}(t)$ and  $w_{l,\alpha}(t)$ is created for each driver.
@@
 \paragraph{4. Temporal Anticipation:}
-The HDM models an experienced driver's anticipation by having driver's assume that their acceleration is constant over the reaction time $T_r$.
+The HDM models an experienced driver's anticipation by having drivers assume that their own acceleration is constant over the reaction time $T_r$.
 This assumption allows a driver to predict what their own speed will be $T_r$ seconds in the future.
@@
 Moreover, in the HDM, a driver assumes that the velocity of leading vehicle is constant over next $T_r$ seconds.
-This allows the driver to make predictions of the distance to and speed of the leading vehicle, as shown in Eq (21, 22) respectively, where $\Delta v_\alpha^\text{est}=v_\alpha-v_l^\text{est}$.
+This allows the driver to anticipate of the distance to and speed of the leading vehicle, as shown in Eq (21, 22) respectively, where $\Delta v_\alpha^\text{est}=v_\alpha-v_l^\text{est}$.
 \begin{mymathbox}[ams gather, title=Temporal Anticipation Equations,colframe=blue!30!black]
@@
 \paragraph{5. Multi-Vehicle Anticipation:}
-The HDM models allows a driver to be aware multiple vehicles ahead.
-The acceleration function (shown in Eq(23)) is broken up into two terms: a free acceleration term that representing the acceleration on an open road, and an interaction term that models the effect of the driver's awareness of multiple vehicles in front of it on the driver's acceleration.
-The free acceleration and interaction terms directly follow from adapting the IDM acceleration function to include the distances to and speeds of multiple leading cars.
-The coefficient of the interaction term, $c_{\text{IDM}}$, is used to keep the interaction term from blowing up in the negative direction as $n_a$ increases.
+The HDM allows a driver to be aware multiple vehicles ahead.
+The acceleration function (shown in Eq(23)) is broken up into two terms: a free acceleration term that representing the acceleration on an open road, and an interaction term that models the effect, on the acceleration, of a driver's awareness of multiple leading vehicles.
+The free acceleration and interaction terms directly follow from adapting the IDM acceleration function to include the distances to and speeds of multiple leading vehicles.
+The coefficient of the interaction term $c_{\text{IDM}}$ is used to keep the interaction term from blowing up in the negative direction as $n_a$, the number of vehicles anticipated, increases.
 \begin{mymathbox}[ams gather, title=Multi-Vehicle Anticipation Equations,colframe=blue!30!black]

### 2016-12-12 project_paper/paper.tex
added conclusion

@@
 
-\section{TODO}
-fuck this to do list, we've done enough. \\
-develop a cost function for emissions/energy usage and optimize over that \\
-can we find the parameters that will cause a phantom traffic jam?
-
-\section{Discussion}
-
 \section{Conclusion}
+The models used within this paper necessarily make many simplifying assumptions, many of which vary somewhat between models.
+Because of this, it was often difficult to maintain stability of the systems in blended models, as one model may not fully satisfy the assumptions of the other under all parameters.
+However, within these restrictions, we were still able to construct a series of simulations with stable parameter values and model the reactions of the systems in reasonable ways.
+We find that in general ACC-controllers are better than human drivers at adapting to and dissipating traffic disturbances and that introducing more self-driving cars in a simplified scenario such as this is likely to improve traffic patterns by reducing the propagation and amplification of traffic waves.
 

### 2016-12-12 project_paper/paper.tex
added some editing etc

@@
 This is mainly due to the inherent random movements of cars, the number of individually moving cars and the many and varying scenarios that can be analyzed.
-In this project, we present two main models, the `Intelligent Driver Model' and the `Human Driver Model' and numerically simulate a number of test cases allowing us to study the underlying dynamics of this system.
+In this project we present two main models: the `Intelligent Driver Model' and the `Human Driver Model'.
+We numerically simulate a number of test cases, allowing us to study the underlying dynamics of this system.
 We focus on modeling the cars using a differential setting where one car's motion is dependent on the car that is in front of it (and sometimes dependent on multiple cars in front of it).
-We will then extend the scope of this analysis and introduce a number of varying factors into the model to analyze the motion of cars along a straight road, analyze the cars along a circular track and finally include a blend of the intelligent driving cars (comparable to automated cruise control autonomous cars) and cars with the human controlled vehicles.
+We will then extend the scope of this analysis and introduce a number of varying factors into the model to analyze the motion of cars along a straight road, analyze the cars along a circular track and finally include a blend of the intelligent driving cars (comparable to automated cruise control autonomous cars) and the human controlled vehicles.
 These autonomous cars have more precise, and faster reacting velocity controllers and the inclusion of vehicles such as these in the model is expected to smooth the flow of traffic and ultimately increase the density of traffic flow.
@@
 \paragraph{} In the microscopic perspective, individual cars are modeled as particles that move according to a relationship with the particle(s) in front of them.
-Time-continuous microscopic models, commonly referred to as car-following models, model the characteristics of individual vehicles such as acceleration, velocity, position using a systems of coupled ordinary differential equations (ODEs).
-Car-following models are constained by assumptions that mimic realistic driving behavior such as a vehicle wants to maintain a safe driving gap between itself and vehicle in front of it (commonly called the \textit{leading vehicle}), a vehicle has a certain desired velocity, or there is a maximum comfortable acceleration \cite[chapter~10]{treiber_kesting_2013}.
-In the section, we consider a few variations of the Intelligent Driver Model (IDM), as well as Human Driver Model (HDM).
+Time-continuous microscopic models, commonly referred to as car-following models, model the characteristics of individual vehicles using systems of coupled ordinary differential equations (ODEs).
+These characteristics include the acceleration, velocity and position of the cars.
+Car following models are constrained by assumptions \cite[chapter~10]{treiber_kesting_2013} that attempt to mimic realistic driving behavior such as:
+\begin{itemize}
+\item A vehicle wants to maintain a safe driving gap between itself and vehicle in front of it (commonly called the \textit{leading vehicle}).
+\item A vehicle has a certain desired velocity (e.g. the speed limit).
+\item There is a maximum comfortable acceleration (or maximum constraint of acceleration from the physical limits of the car).
+\end{itemize}
 %We chose the IDM over other car-following models such as Gipps Model because the IDM has direct extensions to modeling human driving behavior using the HDM.
-We chose the IDM over other car-following models such as the Gipps Model or Newell's due to the oversimplification of some acceleration assumptions that the alternative models make\cite[chapter~11]{treiber_kesting_2013}.
+We chose the IDM in this project over other car-following models such as the Gipps Model or Newell's due to the oversimplification of some acceleration assumptions that the alternative models make \cite[chapter~11]{treiber_kesting_2013}.
 The result is that the IDM provides the simplest implementation of plausible driving characteristics.
-\subsection{Intelligent Driver Models}
-In this section, we model streams of semi-autonomous vehicles that display the following driving characteristics:
+We therefore model streams of semi-autonomous vehicles that display the following driving characteristics:
 \begin{itemize}
-  \item No estimation errors: each vehicle is aware of its own speed, the gap between itself the car in front of it, and the speed of the car in front of it.
-  \item Zero response time: there is no delay between observing a stimulus (i.e. change in gap between itself and vehicle ahead of it or change the vehicle ahead's speed) and responding to it.
-  \item Constant attention: the vehicle is constantly looking for changes in speed or gap
+  \item No estimation errors: each vehicle is able to measure exactly its own speed, the gap between itself the car and in front of it, and the speed of the car in front of it.
+  \item Zero response time: there is no delay between observing a stimulus (i.e. change in gap between itself and vehicle ahead of it or change the leading vehicle's speed) and responding to it.
+  % \item Constant attention: the vehicle is constantly looking for changes in speed or gap.
   \item Each vehicle is only aware of itself and the vehicle directly in front of it.  I.e. it does not account for changes in speed or gap of a vehicle two cars ahead of it.
-  \item Each vehicle is not aware of vehicles behind it or in other lanes
+  % \item Each vehicle is not aware of vehicles behind it or in other lanes
   \item Vehicles are not aware other indicators such as brake lights, horns, headlight flashes, etc.
 \end{itemize}
+
+\subsection{Intelligent Driver Models}
 \subsubsection{Intelligent Driver Model (IDM)}
@@
 The IDM is a simple, accident-free model that can model traffic systems under all realistic traffic conditions.
-The acceleration function, shown in Eq (1), of a particular vehicle $\alpha$ is a function the distance between itself and the leading vehicle ($s_\alpha$, referred to as the gap), the desired speed $v_0$, and the speed of the leading vehicle $v_l$.
+The acceleration function, of a particular vehicle $\alpha$ is a function of the distance between itself and its leading vehicle ($s_\alpha$), the vehicle's desired speed ($v_0$) and the speed of the leading vehicle ($v_l$).
+This is shown in Eq (1) and we further note that $s_\alpha$ is referred to as `the gap'.
 
@@
 \paragraph{}
-The first term in Eq (1) is the \textit{free acceleration}\---the acceleration that a vehicle would take if it were on a wide open road, without considerations of desired speed or gap.
-The second term in Eq (1) is the effect on the vehicle's acceleration as it approaches its desired speed.
-The third term in Eq(1) is the effect, on the acceleration, of relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance.
-The desired gap distance given by Eq (4) is the sum of a desired safe distance $s_0+v_\alpha T$ and a dynamical term $\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$ representing the intelligent vehicle's braking strategy \cite{intelligent_driver_model}.
+The first term in Eq (1) is the \textit{free acceleration} term.
+This is the acceleration that a vehicle would take if it were on a wide open road and had no considerations of desired speed or gap.
+The second term in Eq (1) is the damping effect on the vehicle's acceleration as it approaches its desired speed.
+The third term in Eq(1) is the effect of the relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance on the car's acceleration.
+The desired gap distance, given by Eq (4), is the sum of a desired safe distance $s_0+v_\alpha T$ and a dynamical term ($\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$) representing the intelligent vehicle's braking strategy \cite{intelligent_driver_model}.
 
@@
 \paragraph{}
-The Improved Intelligent Driver Model is an IDM model with an enhanced acceleration function that aims to correct two unrealistic qualities of the IDM when a vehicle's speed is close to the desired speed:
+The Improved Intelligent Driver Model is an IDM model with an enhanced acceleration function that aims to correct two unrealistic qualities of the IDM when a vehicle's speed is close to the desired speed.
+This model makes the following improvements:
 \begin{enumerate}
@@
 The IIDM governing equations seek to improve the unrealistic qualities of the IDM model.
-Eq (8) is the free acceleration equation, which limits the deceleration's magnitude when $v_\alpha>v_0$ to be at most $b$.
-This prevents vehicles from slamming on the brakes as soon as they go faster than the desire velocity.
-To improve the acceleration function's behavior near the desired speed, the IIDM sets the acceleration differently depending on whether the speed and gap above or below their respective desired values.
+Eq (8) is the free acceleration equation, which limits the deceleration's magnitude when $v_\alpha>v_0$ to be at most $b$ (the maximum deceleration that a vehicle may assume).
+This prevents vehicles from slamming on the brakes as soon as they go faster than the desired velocity.
+To improve the acceleration function's behavior near the desired speed, the IIDM sets the acceleration differently depending on whether the speed and gap are above or below their respective desired values.
 When $v_\alpha < v_0$, the IIDM's acceleration function (Eq (5) and Eq (6)) ensures that the actual gap is no larger than the minimum safe gap $s_\alpha = s_0+vT$.
@@
 The ACC is an improvement on the IIDM model that is designed so that a vehicle can handle discontinuous changes in the gap between itself and the leading vehicle or discontinuous changes in the speed of the leading vehicle.
-One scenario such discontinuous changes would occur is when the leading vehicle either enters or exits the lane.
+One scenario where such discontinuous changes would occur is when the leading vehicle either enters or exits the lane.
 Since the IDM based models are only meant for controlling the acceleration of a vehicle within a single lane, this scenario is equivalent to the leading vehicle simply vanishing into or appearing out of thin air.
 In the IDM and IIDM models, vehicles overreact if the gap is suddenly decreased because the models are designed to avoid accidents under all circumstances.
-However, a vehicle slamming on the breaks when the gap is decreased is unrealistic and can even be dangerous since most human drivers do not anticipate sudden changes in the acceleration of the vehicle in front of it.
-It is important note that the ACC model is not guaranteed to be accident-free; the driving style is less defensive by design.
+However, a vehicle slamming on the breaks when the gap is decreased is unrealistic and can even be dangerous since most human and autonomous drivers do not anticipate sudden changes in the acceleration of the vehicle in front of it.
+It is important to note that the ACC model is not guaranteed to be accident-free; the driving style is less defensive by design.
 \paragraph{}
-To attenuate the semi-autonomous vehicle's reaction to sudden decreases in the gap distance, vehicles in the ACC model act according to a constant-acceleration heuristic, rather than making decisions based off the worst-case scenario.
+To attenuate the semi-autonomous vehicle's reaction to sudden decreases in the gap distance, vehicles in the ACC model act according to a constant-acceleration heuristic.
+This is done rather than making decisions based off the worst-case scenario.
 The constant-acceleration heuristic (CAH) assumes that vehicles will have constant acceleration in the near future, no safe distance or time gap is necessary at any time, and that vehicles will react instantaneously.
@@
 One important distinction between the ACC and the previous intelligent driving models is that the ACC uses the leading vehicle's acceleration as an input, in addition to the leading vehicle's speed and gap.
-It is important to note that the CAH acceleration is only used as indicator of whether the IIDM acceleration is overreacting to the situation.
+It is important to note that the CAH acceleration is only used as an indicator of whether the IIDM acceleration is overreacting to the situation.
 Hence, the ACC model also calculates the acceleration function from the IIDM model.
-The acceleration function for the ACC model is given by Eq (9), where $c$ is the "coolness parameter", which is used to tune the ACC model's behavior: $c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model.
+The acceleration function for the ACC model is given by Eq (9), where $c$ is the `coolness parameter', which is used to tune the ACC model's behavior (i.e.
+$c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model).
 If the IIDM acceleration is greater than or equal to the CAH acceleration, then the ACC model uses the IIDM acceleration, since the IIDM model is accident-free and thus guaranteed to be safe.
@@
 \begin{enumerate}
-  \item The\textit{ reaction times} are no longer zero.
-Humans have delayed reactions.
-The HDM requires a reaction time parameter $T_r$, which is the delay in response to changes in the system.
-Hence, the the HDM is governed by a system of time-delayed differential equations.
+  \item The\textit{ reaction times} are no longer zero as humans have delayed reactions.
+The HDM requires a reaction time parameter $T_r$, which is the delay in the driver's response to changes in the system.
+The the HDM is therefore governed by a system of time-delayed differential equations.
   \item For a given driver, the exact distance to and speed of the leading vehicle are not known.
 The driver estimates these quantities with inherent \textit{estimation error}.
-The HDM assumes that the driver knows their own velocity, since the vehicle has a speedometer.
-  \item A driver does not update their acceleration exactly to the value given by the acceleration function.
-This is because a driver controls the car by operating the gas and brake pedals, which leads to imprecise accelerations and induces \textit{driver control errors}.
-  \item Experienced drivers \textit{anticipate} what will happen in the near future and use this anticipation to inform their own behavior.
-  \item Human drivers are able to look ahead multiple vehicles in a single lane of traffic.
+The HDM assumes that the drivers know their own velocities (this assumption is reasonable as the vehicle has a speedometer).
+  \item A driver does not update his acceleration exactly to the value given by the acceleration function.
+This is because a driver controls the car by operating the gas and brake pedals.
+This leads to imprecise accelerations and induces \textit{driver control errors}.
+  \item Experienced drivers \textit{anticipate} what will happen in the near future and use this anticipation to inform their own behaviors.
+  \item Human drivers are able to see multiple vehicles ahead in a single lane of traffic.
 For example, a human driver often can see brake lights of a vehicle three cars ahead on a highway and will react to the brake lights in the distance.
-The HDM models this \textit{multi-vehicle anticipation} allowing the human drivers to update their acceleration based multiple cars ahead of them.
+The HDM models this \textit{multi-vehicle anticipation} by allowing the human drivers to update their acceleration based upon perceived changes that are occurring to multiple cars ahead of them.
 \end{enumerate}
@@
 The HDM incorporates human reaction time by introducing a new reaction time parameter $T_r$.
-Vehicles in the model then update their acceleration based delayed information (i.e.
-the state of the system $T_r$ seconds before) as shown in Eq (13).
-Eq (13) cannot be solved analytically, but $u(t-T_r)$ can be approximated numerically by linear interpolation using the scheme in Eq (14), where $u$ is a placeholder for any of $s_\alpha,v_\alpha,\Delta v_\alpha,v_l$.
+Eq (13) shows that vehicles in the model update their acceleration based on delayed information (i.e.
+the state of the system $T_r$ seconds before).
+It has been shown that Eq (13) cannot be solved analytically, but $u(t-T_r)$ can be approximated numerically by linear interpolation using the scheme in Eq (14), where $u$ is a placeholder for any of $s_\alpha,v_\alpha,\Delta v_\alpha,v_l$.
 In this numerical scheme, $\Delta t$ is the update time and $i=\frac{t}{\Delta t}$ is the update step number.
@@
 \paragraph{2. Estimation Errors:}
-The HDM addresses a driver's error in estimating the distance to and speed of the vehicle in front of it by adding Gaussian noise to the true value of the estimand.
-However, it is likely a particular driver's estimation error will be correlated over time.
+The HDM addresses a driver's error in estimating the distance to and the speed of the vehicle in front of it by adding Gaussian noise to the true value of the estimand.
+However, it is likely that a particular driver's estimation error will be correlated over time (e.g.
+if you perceive that you are a certain distance from the car in front, this estimate will inform your belief at one time step in the future).
 Hence, for each driver $\alpha$, the HDM creates two stochastic processes ($w_{s,\alpha}(t),w_{l,\alpha}(t)$) to represent the driver's error in estimating the distance to and speed of the leading vehicle at each time step, respectively.
@@
 Moreover, each stochastic process uses the autocorrelation function given by Eq (15).
-Each stochastic process is created approximating a Wiener Process using the numerical scheme given in Eq (16), where $i=\frac{t}{\Delta t}$ is the update step number and $\eta_i \sim N(0,1)$.
+These processes are created by approximating a Wiener Process using the numerical scheme given in Eq (16), where $i=\frac{t}{\Delta t}$ is the update step number and $\eta_i \sim N(0,1)$.
 Using these stochastic processes, the estimation errors (or log errors) of distance to and speed of the leading vehicle are given by Eq (17, 18), respectively, where $s_\alpha^\text{est}$ and $v_l^\text{est}$ are the estimates.
@@
 This prediction is given by Eq (20).
-Moreover, in the HDM, a driver assumes that the velocity of leading vehicle is constant over next $T_r$ seconds.
-This allows the driver to anticipate of the distance to and speed of the leading vehicle, as shown in Eq (21, 22) respectively, where $\Delta v_\alpha^\text{est}=v_\alpha-v_l^\text{est}$.
+Moreover, in the HDM, a driver assumes that the velocity of the leading vehicle is constant over next $T_r$ seconds.
+This allows the driver to anticipate the distance to and speed of the leading vehicle, as shown in Eq (21, 22) respectively, where $\Delta v_\alpha^\text{est}=v_\alpha-v_l^\text{est}$.
 \begin{mymathbox}[ams gather, title=Temporal Anticipation Equations,colframe=blue!30!black]
@@
 \paragraph{5. Multi-Vehicle Anticipation:}
-The HDM allows a driver to be aware multiple vehicles ahead.
-The acceleration function (shown in Eq(23)) is broken up into two terms: a free acceleration term that representing the acceleration on an open road, and an interaction term that models the effect, on the acceleration, of a driver's awareness of multiple leading vehicles.
+The HDM allows a driver to be aware of multiple vehicles ahead of it.
+The acceleration function (shown in Eq(23)) is broken up into two terms: a free acceleration term that represents the acceleration on an open road and an interaction term that models the effect, on the acceleration, of a driver's awareness of multiple leading vehicles.
 The free acceleration and interaction terms directly follow from adapting the IDM acceleration function to include the distances to and speeds of multiple leading vehicles.
@@
 We explored scenarios of all IDM cars and all HDM cars.
-Under these conditions, we see the IDM models generate more `phantom` traffic waves than the HDM models, especially for larger numbers of HDM car lookahead.
-This is because the human drivers tend to appear more conservative in a situation like this, with no external perturbations, as the multi-car lookahead prevents them from getting overly close to the car directly in front of them.
+Under these conditions, we see the IDM models generate more `phantom' traffic waves than the HDM models, especially for larger numbers of HDM car lookahead.
+This is because the human drivers tend to appear more conservative in a situation like this (i.e.
+with no external perturbations) as the multi-car lookahead prevents them from getting overly close to the car directly in front of them.
 
@@
 \end{figure}
-\paragraph{}For our purposes, we desired to use higher order methods such that we could use larger timesteps and hopefully decrease running time of our programs while still maintaining stability in the differential equation systems.
+\paragraph{}For our purposes, we used higher order methods as we were then able use larger timesteps and hopefully decrease the running time of our programs while still maintaining stability in the differential equation systems.
 To display the effectiveness of these higher order methods, we ran one of our simple models for a variety of time steps over the same total time with the different solvers and noted the number of time steps at which each method stabilized.
@@
 \section{Optimization of Traffic Systems}
-\paragraph{}We set up a number of optimization programs to attempt to search for the optimal values of some of the many parameters for the system as a whole and for individual drivers for a number of cost functions, including maximizing displacement and maximizing comfort (minimizing jerk).
+\paragraph{}We set up a number of optimization programs to attempt to search for the optimal values of some of the many parameters for the system as a whole and for individual drivers for a number of cost functions, including maximizing displacement and maximizing comfort (minimizing jerk factor).
 The initial trials for these systems were run on IDM models, and the results are largely what one would suspect.
@@
   \fbox{\includegraphics[width=0.95\textwidth]{optimise_comfort.png}}
-  \caption{System of IDM cars on a circular track optimizing for maximum 'comfort' (minimum jerk). As one might expect, they choose to go rather slowly and thereby achieve a very stable velocity.}
+  \caption{System of IDM cars on a circular track optimizing for maximum `comfort' (minimum jerk). As one might expect, they choose to go rather slowly and thereby achieve a very stable velocity.}
 \end{figure}
@@
 \\
-Again, we find that in general maximizing displacement for a single car leads to greater displacement for the rest of the system unless the driving behavior is so aggressive that it results in a crash.
+Again, we find that in general, maximizing displacement for a single car leads to greater displacement for the rest of the system unless the driving behavior is so aggressive that it results in a crash.
 Given that we have seen that the Human Driver Model is more likely to propagate perturbations, we had expected that there might be in this model a penalty to other cars in greater volatility of velocity, i.e.

### 2016-12-13 project_paper/paper.tex
final edits and added references

@@
 \paragraph{}There are generally two ways to model traffic: from a microscopic and from a macroscopic perspective \cite[Chapter~34]{microscopic_modelling}.
-In the macroscopic perspective, traffic is viewed as a fluid or gas with a given maximum density moving according to the laws of conservation of mass.
+In the macroscopic models, traffic is viewed as a fluid or gas (with a given maximum density) and it moves according to the laws of conservation of mass.
 Because we'd like to interweave different models for specific cars, these models would be difficult to blend in a way that is easily interpretable.
@@
 \section{Microscopic Perspective of Single Lane Traffic Interactions}
-\paragraph{} In the microscopic perspective, individual cars are modeled as particles that move according to a relationship with the particle(s) in front of them.
+\paragraph{} In the microscopic models, individual cars are modeled as particles that move according to a relationship with the particle(s) in front of them.
 Time-continuous microscopic models, commonly referred to as car-following models, model the characteristics of individual vehicles using systems of coupled ordinary differential equations (ODEs).
@@
   % \item Each vehicle is not aware of vehicles behind it or in other lanes
-  \item Vehicles are not aware other indicators such as brake lights, horns, headlight flashes, etc.
+  \item Vehicles are not aware of other indicators such as brake lights, horns, headlight flashes, etc.
 \end{itemize}
@@
 This is shown in Eq (1) and we further note that $s_\alpha$ is referred to as `the gap'.
+\cite[chapter~11]{treiber_kesting_2013}
 
@@
 \paragraph{}
-The IIDM governing equations seek to improve the unrealistic qualities of the IDM model.
+The IIDM governing equations seek to improve the unrealistic qualities of the IDM model \cite{intelligent_driver_model}.
 Eq (8) is the free acceleration equation, which limits the deceleration's magnitude when $v_\alpha>v_0$ to be at most $b$ (the maximum deceleration that a vehicle may assume).
@@
 \paragraph{}
-The ACC is an improvement on the IIDM model that is designed so that a vehicle can handle discontinuous changes in the gap between itself and the leading vehicle or discontinuous changes in the speed of the leading vehicle.
+The ACC is an improvement on the IIDM model that is designed so that a vehicle can handle discontinuous changes in the gap between itself and the leading vehicle or discontinuous changes in the measured speed of the leading vehicle \cite[chapter~11]{treiber_kesting_2013}.
 One scenario where such discontinuous changes would occur is when the leading vehicle either enters or exits the lane.
@@
 It is important to note that the ACC model is not guaranteed to be accident-free; the driving style is less defensive by design.
+\cite{loos_platzer_nistor_2011}
 \paragraph{}
@@
 This is done rather than making decisions based off the worst-case scenario.
-The constant-acceleration heuristic (CAH) assumes that vehicles will have constant acceleration in the near future, no safe distance or time gap is necessary at any time, and that vehicles will react instantaneously.
+The constant-acceleration heuristic (CAH) assumes that vehicles will have constant acceleration in the near future, no safe distance or time gap is necessary at any time, and that vehicles will react instantaneously.\cite[chapter~11]{treiber_kesting_2013}
 \begin{mymathbox}[ams gather, title=ACC Governing Functions, colframe=blue!30!black]
@@
 The acceleration function for the ACC model is given by Eq (9), where $c$ is the `coolness parameter', which is used to tune the ACC model's behavior (i.e.
-$c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model).
+$c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model) \cite[chapter~11]{treiber_kesting_2013}.
 If the IIDM acceleration is greater than or equal to the CAH acceleration, then the ACC model uses the IIDM acceleration, since the IIDM model is accident-free and thus guaranteed to be safe.
@@
 \paragraph{}
-The Human Driver Model (HDM) is an extension of the IDM model that models human driving behavior by making five important changes to the IDM:
+The Human Driver Model (HDM) is an extension of the IDM model that models human driving behavior by making five important changes to the IDM \cite[chapter~12]{treiber_kesting_2013}:
 \begin{enumerate}
   \item The\textit{ reaction times} are no longer zero as humans have delayed reactions.
-The HDM requires a reaction time parameter $T_r$, which is the delay in the driver's response to changes in the system.
+The HDM requires a reaction time parameter ($T_r$), which is the delay in the driver's response to changes in the system.
 The the HDM is therefore governed by a system of time-delayed differential equations.
@@
 The driver estimates these quantities with inherent \textit{estimation error}.
-The HDM assumes that the drivers know their own velocities (this assumption is reasonable as the vehicle has a speedometer).
+The HDM assumes that the drivers know their own velocities (this assumption is reasonable as vehicles have speedometers).
   \item A driver does not update his acceleration exactly to the value given by the acceleration function.
-This is because a driver controls the car by operating the gas and brake pedals.
+This is because a driver controls the car by operating the gas and brake pedals and there is an inherent error in this input.
 This leads to imprecise accelerations and induces \textit{driver control errors}.
@@
 However, it is likely that a particular driver's estimation error will be correlated over time (e.g.
-if you perceive that you are a certain distance from the car in front, this estimate will inform your belief at one time step in the future).
+if you perceive that you are a certain distance from the car in front, this estimate will inform your belief at one time step in the future) \cite{human_driver_model}.
 Hence, for each driver $\alpha$, the HDM creates two stochastic processes ($w_{s,\alpha}(t),w_{l,\alpha}(t)$) to represent the driver's error in estimating the distance to and speed of the leading vehicle at each time step, respectively.
@@
 The free acceleration and interaction terms directly follow from adapting the IDM acceleration function to include the distances to and speeds of multiple leading vehicles.
-The coefficient of the interaction term $c_{\text{IDM}}$ is used to keep the interaction term from blowing up in the negative direction as $n_a$, the number of vehicles anticipated, increases.
+The coefficient of the interaction term $c_{\text{IDM}}$ is used to keep the interaction term from blowing up in the negative direction as $n_a$, the number of vehicles anticipated, increases \cite[chapter~12]{treiber_kesting_2013}.
 \begin{mymathbox}[ams gather, title=Multi-Vehicle Anticipation Equations,colframe=blue!30!black]
@@
 \paragraph{}To generate numerical simulations of Traffic Systems, we first consider the case of 50 cars on a circular track.
-We explored scenarios of all IDM cars and all HDM cars.
+We explored scenarios of consisting purely of IDM cars or of HDM cars.
 Under these conditions, we see the IDM models generate more `phantom' traffic waves than the HDM models, especially for larger numbers of HDM car lookahead.
-This is because the human drivers tend to appear more conservative in a situation like this (i.e.
+This is because the human drivers tend to be more more conservative in a situation like this (i.e.
 with no external perturbations) as the multi-car lookahead prevents them from getting overly close to the car directly in front of them.
@@
 For large enough time steps, this shows us error decreasing on a log scale for each of our higher order methods.
-However, for small time steps, the `error` for 4th and 5th order methods converges, as the \texttt{odeint} solver presumably decides that 4th order is sufficiently accurate at these time steps and stops using higher order methods.\\
+However, for small time steps, the `error' for 4th and 5th order methods converges, as the \texttt{odeint} solver presumably decides that 4th order is sufficiently accurate at these time steps and stops using higher order methods.\\
 \begin{figure}[H]
@@
 
-For HDM models, the presence or lack thereof of string stability depends on a number of parameters: number of leading cars being considered, time step ($\Delta t$), and reaction time ($T_r$).
+For HDM models, the presence or lack thereof of string stability depends on a number of parameters: number of leading cars being considered, time step ($\Delta t$), and reaction time ($T_r$) \cite[chapter~15]{treiber_kesting_2013}.
 In general, string stability increases with number of leading cars being considered, although the marginal contribution of each car decreases to the point where there is not much additional benefit beyond 5 cars.
@@
 \section{Optimization of Traffic Systems}
-\paragraph{}We set up a number of optimization programs to attempt to search for the optimal values of some of the many parameters for the system as a whole and for individual drivers for a number of cost functions, including maximizing displacement and maximizing comfort (minimizing jerk factor).
+We set up a number of optimization programs to attempt to search for the optimal values of some of the many parameters for the system as a whole and for individual drivers for a number of cost functions, including maximizing displacement and maximizing comfort (minimizing jerk factor\footnote{Jerk factor is a combination of a car's acceleration and the derivative of acceleration.
+The aim is to minimize any fast changes to the car's acceleration.}) \cite[chapter~21]{treiber_kesting_2013}.
 The initial trials for these systems were run on IDM models, and the results are largely what one would suspect.
@@
 
-\paragraph{}We then attempted to optimize behavior for a single driver and were interested to see what implications this would have for the system as a whole.
+\par
+We then attempted to optimize behavior for a single driver and were interested to see what implications this would have for the system as a whole.
 However, the following behaviors in the IDM are such that a single driver can tailgate and maximize target speed basically as much as we will allow without ever disrupting the rest of the system.
@@
 In fact, it seems to be true generally of these systems that unless the aggressive driving behavior of a car causes other cars to collide behind it, the greater distance achieved by one car inevitably results in other cars also achieving a greater distance/average velocity.
-\\
 
-\paragraph{}Next, we attempted to replicate these experiments with the more complicated Human Driver Model.
+\par
+Next, we attempted to replicate these experiments with the more complicated Human Driver Model.
 However, due to the highly piecewise nature of the functions we are attempting to optimize over, we were unable to use traditional minimization solvers.
-Instead, we discretized the parameter space we were interested in exploring and searched for the set of parameters which would minimize our target cost function.
-\\
-Again, we find that in general, maximizing displacement for a single car leads to greater displacement for the rest of the system unless the driving behavior is so aggressive that it results in a crash.
-Given that we have seen that the Human Driver Model is more likely to propagate perturbations, we had expected that there might be in this model a penalty to other cars in greater volatility of velocity, i.e.
+Instead, we discretized the parameter space that we were interested in exploring and searched for the set of parameters which would minimize our target cost function.
+\par
+Again, we find that in general, maximizing displacement for a single car leads to a greater displacement for the rest of the system unless the driving behavior is so aggressive that it results in a crash.
+Given that we have seen that the Human Driver Model is more likely to propagate perturbations to the system, we had expected that in this model there might be a penalty to other cars in greater volatility of velocity, i.e.
 a less pleasant ride, but this effect was also not observed in the data.
@@
   \fbox{\includegraphics[width=0.95\textwidth]{optimize_t.png}}
-  \caption{Changing Safe Following Time: Changes are minimal but overall distance increases for both the car in question and the system of cars with smaller.}
+  \caption{Changing Safe Following Time: Changes are minimal but overall distance increases for both the car in question and the system of cars with smaller safe following times.}
 \end{figure}

### 2016-12-13 project_paper/paper.tex
tiny edits

@@
 \paragraph{}
-The modeling of traffic at traffic lights, on highways, in dense traffic streams, with varying situations etc can often present a problem that is too complicated to solve analytically.
-This is mainly due to the inherent random movements of cars, the number of individually moving cars and the many and varying scenarios that can be analyzed.
+The modeling of traffic in varying situations (at traffic lights, on highways, in dense traffic streams, etc.) can often present a problem that is too complicated to solve analytically.
+This is mainly due to the inherent random movements of cars, the number of individually moving cars ,and the many and varying scenarios that can be analyzed.
 In this project we present two main models: the `Intelligent Driver Model' and the `Human Driver Model'.
@@
 We will then extend the scope of this analysis and introduce a number of varying factors into the model to analyze the motion of cars along a straight road, analyze the cars along a circular track and finally include a blend of the intelligent driving cars (comparable to automated cruise control autonomous cars) and the human controlled vehicles.
-These autonomous cars have more precise, and faster reacting velocity controllers and the inclusion of vehicles such as these in the model is expected to smooth the flow of traffic and ultimately increase the density of traffic flow.
+These autonomous cars have more precise, and faster reacting velocity controllers, and the inclusion of vehicles such as these in the model is expected to smooth the flow of traffic and ultimately increase the density of traffic flow.
 
@@
 \section{Stability and Error Comparison of a Variety of Numerical Integration Schemes}
-Because there is no analytical solution given for our highly-nonlinear problem, to analyze the error of our numerical integration schemes, we compared to the python \texttt{odeint} solver function.
+\paragraph{}Because there is no analytical solution given for our highly-nonlinear problem, to analyze the error of our numerical integration schemes, we compared to the python \texttt{odeint} solver function.
 For large enough time steps, this shows us error decreasing on a log scale for each of our higher order methods.
@@
 
-The concept of string stability refers to whether a perturbation to the beginning of the system will result in a wave that dissipates or is enhanced as it propagates through the system.
+\paragraph{}The concept of string stability refers to whether a perturbation to the beginning of the system will result in a wave that dissipates or is enhanced as it propagates through the system.
 As we have seen in previous experiments that the IDM models are excellent for adjusting to and dissipating traffic perturbations, we expect them to be string stable and indeed they are.
@@
 
-For HDM models, the presence or lack thereof of string stability depends on a number of parameters: number of leading cars being considered, time step ($\Delta t$), and reaction time ($T_r$) \cite[chapter~15]{treiber_kesting_2013}.
+\paragraph{}For HDM models, the presence or lack thereof of string stability depends on a number of parameters: number of leading cars being considered, time step ($\Delta t$), and reaction time ($T_r$) \cite[chapter~15]{treiber_kesting_2013}.
 In general, string stability increases with number of leading cars being considered, although the marginal contribution of each car decreases to the point where there is not much additional benefit beyond 5 cars.
@@
 \section{Optimization of Traffic Systems}
-We set up a number of optimization programs to attempt to search for the optimal values of some of the many parameters for the system as a whole and for individual drivers for a number of cost functions, including maximizing displacement and maximizing comfort (minimizing jerk factor\footnote{Jerk factor is a combination of a car's acceleration and the derivative of acceleration.
+\paragraph{}We set up a number of optimization programs to attempt to search for the optimal values of some of the many parameters for the system as a whole and for individual drivers for a number of cost functions, including maximizing displacement and maximizing comfort (minimizing jerk factor\footnote{Jerk factor is a combination of a car's acceleration and the derivative of acceleration.
 The aim is to minimize any fast changes to the car's acceleration.}) \cite[chapter~21]{treiber_kesting_2013}.
@@
 
-\par
-We then attempted to optimize behavior for a single driver and were interested to see what implications this would have for the system as a whole.
+\paragraph{}We then attempted to optimize behavior for a single driver and were interested to see what implications this would have for the system as a whole.
 However, the following behaviors in the IDM are such that a single driver can tailgate and maximize target speed basically as much as we will allow without ever disrupting the rest of the system.
@@
 
-\par
-Next, we attempted to replicate these experiments with the more complicated Human Driver Model.
+\paragraph{}Next, we attempted to replicate these experiments with the more complicated Human Driver Model.
 However, due to the highly piecewise nature of the functions we are attempting to optimize over, we were unable to use traditional minimization solvers.
 Instead, we discretized the parameter space that we were interested in exploring and searched for the set of parameters which would minimize our target cost function.
-\par
-Again, we find that in general, maximizing displacement for a single car leads to a greater displacement for the rest of the system unless the driving behavior is so aggressive that it results in a crash.
+
+\paragraph{}Again, we find that in general, maximizing displacement for a single car leads to a greater displacement for the rest of the system unless the driving behavior is so aggressive that it results in a crash.
 Given that we have seen that the Human Driver Model is more likely to propagate perturbations to the system, we had expected that in this model there might be a penalty to other cars in greater volatility of velocity, i.e.
@@
 \section{Conclusion}
-The models used within this paper necessarily make many simplifying assumptions, many of which vary somewhat between models.
+\paragraph{}The models used within this paper necessarily make many simplifying assumptions, many of which vary somewhat between models.
 Because of this, it was often difficult to maintain stability of the systems in blended models, as one model may not fully satisfy the assumptions of the other under all parameters.

### 2016-12-14 project_paper/paper.tex
Final edits

@@
 
-\title{AM205 Project - Traffic Flow Modelling\\
+\title{AM205 Project - Traffic Flow Modeling\\
 \Large{Microscopic Models of Traffic, Comparing the Intelligent Driver Model (with Adaptive Cruise Control) with the Human Driver Model}}
@@
 The modeling of traffic in varying situations (at traffic lights, on highways, in dense traffic streams, etc.) can often present a problem that is too complicated to solve analytically.
-This is mainly due to the inherent random movements of cars, the number of individually moving cars ,and the many and varying scenarios that can be analyzed.
-In this project we present two main models: the `Intelligent Driver Model' and the `Human Driver Model'.
+This is mainly due to the inherent stochasticity in the movements of cars and the large number of individually moving cars.
+In this project we present two main models, the `Intelligent Driver Model' (IDM) and the `Human Driver Model' (HDM), for modeling traffic flow within a single lane.
 We numerically simulate a number of test cases, allowing us to study the underlying dynamics of this system.
-We focus on modeling the cars using a differential setting where one car's motion is dependent on the car that is in front of it (and sometimes dependent on multiple cars in front of it).
-We will then extend the scope of this analysis and introduce a number of varying factors into the model to analyze the motion of cars along a straight road, analyze the cars along a circular track and finally include a blend of the intelligent driving cars (comparable to automated cruise control autonomous cars) and the human controlled vehicles.
-These autonomous cars have more precise, and faster reacting velocity controllers, and the inclusion of vehicles such as these in the model is expected to smooth the flow of traffic and ultimately increase the density of traffic flow.
+We focus on modeling the cars using a differential setting where a vehicle's motion is dependent on one or more vehicles front of it.
+Using each of these models, we analyze the motion of a system of vehicles cars along a straight road and along a circular track.
+Finally we create a system composed of a mixture intelligent driving vehicles (comparable to automated cruise control semi-autonomous cars) and the human-controlled vehicles.
+The semi-autonomous vehicles have precise control of their acceleration, instantaneous reactions, and sensors providing accurate estimates of positions and velocities of the vehicles around them.
+We expect that including of vehicles such as these in the model will smooth the flow of traffic and ultimately increase the density of traffic flow.
 
-\paragraph{}There are generally two ways to model traffic: from a microscopic and from a macroscopic perspective \cite[Chapter~34]{microscopic_modelling}.
+\paragraph{}There are two common ways to model traffic: from a microscopic and from a macroscopic perspective.
 In the macroscopic models, traffic is viewed as a fluid or gas (with a given maximum density) and it moves according to the laws of conservation of mass.
-Because we'd like to interweave different models for specific cars, these models would be difficult to blend in a way that is easily interpretable.
+The microscopic perspective considers systems of individual vehicles each acting according to set of governing differential equations\cite[Chapter~34]{microscopic_modelling}.
+In this paper, we consider microscopic models because we are interested in creating systems composed of a mixture of semi-autonomous and human-controlled vehicles.
+It would be difficult to adapt macroscopic models to model such systems in an interpretable way.
 
 \section{Microscopic Perspective of Single Lane Traffic Interactions}
-\paragraph{} In the microscopic models, individual cars are modeled as particles that move according to a relationship with the particle(s) in front of them.
+\paragraph{} In microscopic models, individual cars are modeled as particles that move according to a relationship with the particle(s) in front of them.
 Time-continuous microscopic models, commonly referred to as car-following models, model the characteristics of individual vehicles using systems of coupled ordinary differential equations (ODEs).
-These characteristics include the acceleration, velocity and position of the cars.
-Car following models are constrained by assumptions \cite[chapter~10]{treiber_kesting_2013} that attempt to mimic realistic driving behavior such as:
+These characteristics include the acceleration, velocity, and position of the cars.
+Car-following models are constrained by assumptions \cite[Chapter~10]{treiber_kesting_2013} that attempt to mimic realistic driving behavior such as:
 \begin{itemize}
 \item A vehicle wants to maintain a safe driving gap between itself and vehicle in front of it (commonly called the \textit{leading vehicle}).
-\item A vehicle has a certain desired velocity (e.g. the speed limit).
-\item There is a maximum comfortable acceleration (or maximum constraint of acceleration from the physical limits of the car).
+\item A vehicle has a certain desired velocity\footnote{Note: we use the terms \textit{speed} and \textit{velocity} interchangeably in this paper since in all models we discuss, vehicles can only have nonnegative velocities.} (e.g.
+the speed limit).
+\item There is a maximum acceleration that can be constrained by either the physical limitations of the car or comfort.
 \end{itemize}
-%We chose the IDM over other car-following models such as Gipps Model because the IDM has direct extensions to modeling human driving behavior using the HDM.
-We chose the IDM in this project over other car-following models such as the Gipps Model or Newell's due to the oversimplification of some acceleration assumptions that the alternative models make \cite[chapter~11]{treiber_kesting_2013}.
-The result is that the IDM provides the simplest implementation of plausible driving characteristics.
-We therefore model streams of semi-autonomous vehicles that display the following driving characteristics:
-\begin{itemize}
-  \item No estimation errors: each vehicle is able to measure exactly its own speed, the gap between itself the car and in front of it, and the speed of the car in front of it.
-  \item Zero response time: there is no delay between observing a stimulus (i.e. change in gap between itself and vehicle ahead of it or change the leading vehicle's speed) and responding to it.
-  % \item Constant attention: the vehicle is constantly looking for changes in speed or gap.
-  \item Each vehicle is only aware of itself and the vehicle directly in front of it.  I.e. it does not account for changes in speed or gap of a vehicle two cars ahead of it.
-  % \item Each vehicle is not aware of vehicles behind it or in other lanes
-  \item Vehicles are not aware of other indicators such as brake lights, horns, headlight flashes, etc.
-\end{itemize}
-
+In this project, we chose the Intelligent Driver Model over other car-following models such as the Gipps Model or Newell's due to the oversimplification of some acceleration assumptions that the alternative models make \cite[Chapter~11]{treiber_kesting_2013}.
 \subsection{Intelligent Driver Models}
@@
 The IDM is a simple, accident-free model that can model traffic systems under all realistic traffic conditions.
+The IDM provides a simple implementation for semi-autonomous vehicles that display the following driving characteristics:
+\begin{itemize}
+  \item No estimation errors: each vehicle is able to measure exactly its own speed, the gap between itself the car and in front of it, and the speed of the car in front of it.
+  \item Zero response time: there is no delay between observing a stimulus (i.e. a change in gap between itself and vehicle ahead of it or a change the leading vehicle's speed) and responding to it.
+  \item Each vehicle is only aware of itself and the vehicle directly in front of it; it does not account for changes in the speed or gap of a vehicle two cars ahead of it.
+  \item Vehicles are not aware of other indicators such as brake lights, horns, headlight flashes, etc.
+\end{itemize}
 The acceleration function, of a particular vehicle $\alpha$ is a function of the distance between itself and its leading vehicle ($s_\alpha$), the vehicle's desired speed ($v_0$) and the speed of the leading vehicle ($v_l$).
 This is shown in Eq (1) and we further note that $s_\alpha$ is referred to as `the gap'.
-\cite[chapter~11]{treiber_kesting_2013}
-
+\cite[Chapter~11]{treiber_kesting_2013}
 \begin{mymathbox}[title=IDM Parameters, colframe=blue!30!black]
@@
 The second term in Eq (1) is the damping effect on the vehicle's acceleration as it approaches its desired speed.
-The third term in Eq(1) is the effect of the relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance on the car's acceleration.
-The desired gap distance, given by Eq (4), is the sum of a desired safe distance $s_0+v_\alpha T$ and a dynamical term ($\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$) representing the intelligent vehicle's braking strategy \cite{intelligent_driver_model}.
-
+The third term in Eq (1) is the effect of the relative difference between the desired gap distance $s^*(v_\alpha,\Delta v_\alpha)$ and the current gap distance on the car's acceleration.
+The difference in position and speed of a vehicle $\alpha$ and the leading vehicle are given by Eq (2) and (3).
+The desired gap distance, given by Eq (4), is the sum of a desired safe distance $s_0+v_\alpha T$ and a term representing the intelligent vehicle's braking strategy \cite{intelligent_driver_model} ($\frac{v_\alpha \Delta v_\alpha}{2 \sqrt{ab}}$).
 \subsubsection{Improved Intelligent Driver Model (IIDM)}
 \paragraph{}
-The Improved Intelligent Driver Model is an IDM model with an enhanced acceleration function that aims to correct two unrealistic qualities of the IDM when a vehicle's speed is close to the desired speed.
-This model makes the following improvements:
+The Improved Intelligent Driver Model is an IDM model with an enhanced acceleration function that aims to correct two unrealistic qualities of the IDM that occur when a vehicle's speed is close to the desired speed.
+This model makes the following improvements \cite[Chapter~11]{treiber_kesting_2013}:
 \begin{enumerate}
   \item
-  If a vehicle exceeds the desired speed, the IDM acceleration function will return a very large negative acceleration.
+  If a vehicle exceeds the desired speed, the IDM vehicle's acceleration controller will effectively slam on the brakes.
 This is unrealistic since drivers on a highway tend to oscillate around their desired speed.
@@
   If a vehicle is close to the desired speed, the desired gap $s^*(v_\alpha,\Delta v_\alpha)$ becomes much wider than $s_0 + v_\alpha T$.
-This is unrealistic and may preclude vehicles from reaching the desired velocity $v_0$.
+This may preclude vehicles from reaching the desired velocity $v_0$.
 \end{enumerate}
@@
 When $v_\alpha < v_0$, the IIDM's acceleration function (Eq (5) and Eq (6)) ensures that the actual gap is no larger than the minimum safe gap $s_\alpha = s_0+vT$.
-To maintain the intelligent and comfortable brake method of the IDM, this improved acceleration function mimics the behavior of the IDM acceleration function except for when the actual gap is near the desired gap or when the actual speed is greater than the desired speed.
+To maintain the intelligent and comfortable brake method of the IDM, this improved acceleration function mimics the behavior of the IDM acceleration function except for when the actual gap is near the desired gap or when the actual speed is greater than the desired speed \cite[Chapter~11]{treiber_kesting_2013}.
 \subsubsection{Adaptive Cruise Control Model (ACC)}
 \paragraph{}
-The ACC is an improvement on the IIDM model that is designed so that a vehicle can handle discontinuous changes in the gap between itself and the leading vehicle or discontinuous changes in the measured speed of the leading vehicle \cite[chapter~11]{treiber_kesting_2013}.
+The ACC is an improvement on the IIDM model that is designed so that a vehicle can handle discontinuous changes in the gap between itself and the leading vehicle or discontinuous changes in the measured speed of the leading vehicle \cite[Chapter~11]{treiber_kesting_2013}.
 One scenario where such discontinuous changes would occur is when the leading vehicle either enters or exits the lane.
@@
 In the IDM and IIDM models, vehicles overreact if the gap is suddenly decreased because the models are designed to avoid accidents under all circumstances.
-However, a vehicle slamming on the breaks when the gap is decreased is unrealistic and can even be dangerous since most human and autonomous drivers do not anticipate sudden changes in the acceleration of the vehicle in front of it.
-It is important to note that the ACC model is not guaranteed to be accident-free; the driving style is less defensive by design.
-\cite{loos_platzer_nistor_2011}
+The ACC dampens this response because a vehicle slamming on the brakes when the gap is decreased is unrealistic and can even be dangerous if the system is composed of a mixture of semi-autonomous vehicles and human-controlled vehicles, since most human drivers do not anticipate sudden changes in the acceleration of the vehicle in front of it.
+It is important to note that the ACC model is not guaranteed to be accident-free; the driving style is less defensive by design  \cite[Chapter~11]{treiber_kesting_2013}\cite{loos_platzer_nistor_2011}.
 \paragraph{}
@@
 This is done rather than making decisions based off the worst-case scenario.
-The constant-acceleration heuristic (CAH) assumes that vehicles will have constant acceleration in the near future, no safe distance or time gap is necessary at any time, and that vehicles will react instantaneously.\cite[chapter~11]{treiber_kesting_2013}
+The constant-acceleration heuristic (CAH) assumes that vehicles will have constant acceleration in the near future, no safe distance or time gap is necessary at any time, and that vehicles will react instantaneously \cite[Chapter~11]{treiber_kesting_2013}.
 \begin{mymathbox}[ams gather, title=ACC Governing Functions, colframe=blue!30!black]
@@
 One important distinction between the ACC and the previous intelligent driving models is that the ACC uses the leading vehicle's acceleration as an input, in addition to the leading vehicle's speed and gap.
-It is important to note that the CAH acceleration is only used as an indicator of whether the IIDM acceleration is overreacting to the situation.
+It is important to note that the  acceleration under the constant-acceleration heuristic ($a_{\text{CAH},\alpha}$) is only used as an indicator of whether the IIDM acceleration is overreacting to the situation.
 Hence, the ACC model also calculates the acceleration function from the IIDM model.
-The acceleration function for the ACC model is given by Eq (9), where $c$ is the `coolness parameter', which is used to tune the ACC model's behavior (i.e.
-$c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model) \cite[chapter~11]{treiber_kesting_2013}.
+The acceleration function for the ACC model is given by Eq (9), where $c$ is the `coolness parameter', which is used to tune the ACC model's behavior ($c=0$ is equivalent to a pure IIDM model and $c=1$ is a pure ACC model) \cite[Chapter~11]{treiber_kesting_2013}.
 If the IIDM acceleration is greater than or equal to the CAH acceleration, then the ACC model uses the IIDM acceleration, since the IIDM model is accident-free and thus guaranteed to be safe.
-Otherwise, the AAC model uses a less cautious acceleration function.
+Otherwise, the ACC model uses a less cautious acceleration function.
 \section{Microscopic Perspective of Traffic Interaction: Human Driver Models}
 \paragraph{}
-The Human Driver Model (HDM) is an extension of the IDM model that models human driving behavior by making five important changes to the IDM \cite[chapter~12]{treiber_kesting_2013}:
+The Human Driver Model (HDM) is an extension of the IDM model that models human driving behavior by making five important changes to the IDM \cite[Chapter~12]{treiber_kesting_2013}:
 \begin{enumerate}
-  \item The\textit{ reaction times} are no longer zero as humans have delayed reactions.
+  \item The\textit{ reaction times} are no longer zero because humans have delayed reactions.
 The HDM requires a reaction time parameter ($T_r$), which is the delay in the driver's response to changes in the system.
@@
 The driver estimates these quantities with inherent \textit{estimation error}.
-The HDM assumes that the drivers know their own velocities (this assumption is reasonable as vehicles have speedometers).
+The HDM assumes that the drivers know their own velocities (this assumption is reasonable since vehicles have speedometers).
   \item A driver does not update his acceleration exactly to the value given by the acceleration function.
-This is because a driver controls the car by operating the gas and brake pedals and there is an inherent error in this input.
+This is because a driver controls the car by operating the gas and brake pedals, and there is an inherent error in this input.
 This leads to imprecise accelerations and induces \textit{driver control errors}.
@@
 These processes are created by approximating a Wiener Process using the numerical scheme given in Eq (16), where $i=\frac{t}{\Delta t}$ is the update step number and $\eta_i \sim N(0,1)$.
-Using these stochastic processes, the estimation errors (or log errors) of distance to and speed of the leading vehicle are given by Eq (17, 18), respectively, where $s_\alpha^\text{est}$ and $v_l^\text{est}$ are the estimates.
+Using these stochastic processes, the estimation errors (or log errors) of distance to and speed of the leading vehicle are given by Eq (17, 18), respectively, where $s_\alpha^\text{est}$ and $v_l^\text{est}$ are the estimates \cite[Chapter~12]{treiber_kesting_2013}.
 \begin{mymathbox}[ams gather, title=Estimation Error Equations,colframe=blue!30!black]
@@
 \paragraph{3. Driver Control Errors:}
-The HDM includes driver control errors by adding noise to the acceleration function at each time step.
+The HDM includes driver control errors by adding noise to the acceleration function at each time step \cite[Chapter~12]{treiber_kesting_2013}.
 Specifically, another stochastic process $w_{a,\alpha}(t)$ with the same properties as $w_{s,\alpha}(t)$ and  $w_{l,\alpha}(t)$ is created for each driver.
@@
 This prediction is given by Eq (20).
-Moreover, in the HDM, a driver assumes that the velocity of the leading vehicle is constant over next $T_r$ seconds.
+Moreover, in the HDM, a driver assumes that the velocity of the leading vehicle is constant over next $T_r$ seconds \cite[Chapter~12]{treiber_kesting_2013}.
 This allows the driver to anticipate the distance to and speed of the leading vehicle, as shown in Eq (21, 22) respectively, where $\Delta v_\alpha^\text{est}=v_\alpha-v_l^\text{est}$.
@@
 The HDM allows a driver to be aware of multiple vehicles ahead of it.
-The acceleration function (shown in Eq(23)) is broken up into two terms: a free acceleration term that represents the acceleration on an open road and an interaction term that models the effect, on the acceleration, of a driver's awareness of multiple leading vehicles.
+The acceleration function (shown in Eq (23)) is broken up into two terms: a free acceleration term that represents the acceleration on an open road and an interaction term that models the effect, on the acceleration, of a driver's awareness of multiple leading vehicles.
 The free acceleration and interaction terms directly follow from adapting the IDM acceleration function to include the distances to and speeds of multiple leading vehicles.
-The coefficient of the interaction term $c_{\text{IDM}}$ is used to keep the interaction term from blowing up in the negative direction as $n_a$, the number of vehicles anticipated, increases \cite[chapter~12]{treiber_kesting_2013}.
+The coefficient of the interaction term $c_{\text{IDM}}$ is used to keep the interaction term from blowing up in the negative direction as $n_a$, the number of vehicles anticipated, increases \cite[Chapter~12]{treiber_kesting_2013}.
 \begin{mymathbox}[ams gather, title=Multi-Vehicle Anticipation Equations,colframe=blue!30!black]
@@
 
-\paragraph{}To generate numerical simulations of Traffic Systems, we first consider the case of 50 cars on a circular track.
-We explored scenarios of consisting purely of IDM cars or of HDM cars.
-Under these conditions, we see the IDM models generate more `phantom' traffic waves than the HDM models, especially for larger numbers of HDM car lookahead.
+\paragraph{}To generate numerical simulations of traffic systems, we first considered the case of 50 vehicles on a circular track.
+We explored scenarios where the system is a platoon of either entirely IDM vehicles or entirely HDM vehicles.
+Under these conditions, we see the IDM models generate more `phantom' traffic waves than the HDM models, especially when set larger values for the number of vehicles ahead that a human driver anticipates HDM car lookahead.
 This is because the human drivers tend to be more more conservative in a situation like this (i.e.
@@
 \section{Stability and Error Comparison of a Variety of Numerical Integration Schemes}
-\paragraph{}Because there is no analytical solution given for our highly-nonlinear problem, to analyze the error of our numerical integration schemes, we compared to the python \texttt{odeint} solver function.
-For large enough time steps, this shows us error decreasing on a log scale for each of our higher order methods.
+\paragraph{}Because there is no analytical solution given for our highly-nonlinear problem, we compared several numerical integration schemes to the python \texttt{odeint} solver function to analyze the error of the different numerical schemes.
+For large enough time steps, the error decreases on a log scale for each of our higher order methods.
 However, for small time steps, the `error' for 4th and 5th order methods converges, as the \texttt{odeint} solver presumably decides that 4th order is sufficiently accurate at these time steps and stops using higher order methods.\\
@@
   \fbox{\includegraphics[width=0.95\textwidth]{error.png}}
-  \caption{Error to odeint decreases with accuracy of finite differencing methods.}
+  \caption{Error to \texttt{odeint} decreases with accuracy of finite differencing methods.}
 \end{figure}
@@
 \paragraph{}The concept of string stability refers to whether a perturbation to the beginning of the system will result in a wave that dissipates or is enhanced as it propagates through the system.
-As we have seen in previous experiments that the IDM models are excellent for adjusting to and dissipating traffic perturbations, we expect them to be string stable and indeed they are.
+In previous experiments, we have observed that the IDM models are excellent for adjusting to and dissipating traffic perturbations, so we expect them to be string stable and indeed they are.
 
@@
   \fbox{\includegraphics[width=0.95\textwidth]{displacement_and_velocity_plot_model0_ss.png}}
-  \caption{Plot of cars 1,2,3,6,10, and 20 for IDM. String stable.}
+  \caption{Plot of cars 1, 2, 3, 6, 10, and 20 for IDM. String stable.}
 \end{figure}
 
-\paragraph{}For HDM models, the presence or lack thereof of string stability depends on a number of parameters: number of leading cars being considered, time step ($\Delta t$), and reaction time ($T_r$) \cite[chapter~15]{treiber_kesting_2013}.
+\paragraph{}For HDM models, the presence or lack thereof of string stability depends on a number of parameters: number of leading cars being considered, time step ($\Delta t$), and reaction time ($T_r$) \cite[Chapter~15]{treiber_kesting_2013}.
 In general, string stability increases with number of leading cars being considered, although the marginal contribution of each car decreases to the point where there is not much additional benefit beyond 5 cars.
@@
   \fbox{\includegraphics[width=0.95\textwidth]{HDM_anticipate_2_ss_1000_unstable.png}}
-  \caption{Plot of cars 1,2,3,6,10, and 20 for HDM with two car lookahead.
+  \caption{Plot of cars 1, 2, 3, 6, 10, and 20 for HDM with two car lookahead.
 We see that for parameters $T_r$ = .6 and $\Delta t$ = .2, the system is string unstable, as evidenced by the large wave in car 20's velocity.}
@@
   \fbox{\includegraphics[width=0.95\textwidth]{HDM_anticipate_13002_ss_stable.png}}
-  \caption{Plot of cars 1,2,3,6,10, and 20 with two car lookahead. We see that for parameters $T_r$ = .2 and $\Delta t$ = .15, the system is string stable.}
+  \caption{Plot of cars 1, 2, 3, 6, 10, and 20 with two car lookahead. We see that for parameters $T_r$ = .2 and $\Delta t$ = .15, the system is string stable.}
 \end{figure}
@@
 \paragraph{}We set up a number of optimization programs to attempt to search for the optimal values of some of the many parameters for the system as a whole and for individual drivers for a number of cost functions, including maximizing displacement and maximizing comfort (minimizing jerk factor\footnote{Jerk factor is a combination of a car's acceleration and the derivative of acceleration.
-The aim is to minimize any fast changes to the car's acceleration.}) \cite[chapter~21]{treiber_kesting_2013}.
+The aim is to minimize any fast changes to the car's acceleration.}) \cite[Chapter~21]{treiber_kesting_2013}.
 The initial trials for these systems were run on IDM models, and the results are largely what one would suspect.
@@
 \paragraph{}We then attempted to optimize behavior for a single driver and were interested to see what implications this would have for the system as a whole.
-However, the following behaviors in the IDM are such that a single driver can tailgate and maximize target speed basically as much as we will allow without ever disrupting the rest of the system.
+However, the car-following behaviors in the IDM are such that a single driver can tailgate and maximize target speed basically as much as we will allow without ever disrupting the rest of the system.
 Even when adding some stochasticity to the lead car, we were unable to see the rest of the system suffer from traffic waves due to the behavior of the second car, because that behavior would simply be quickly dissipated by the third car.
 In fact, it seems to be true generally of these systems that unless the aggressive driving behavior of a car causes other cars to collide behind it, the greater distance achieved by one car inevitably results in other cars also achieving a greater distance/average velocity.
-
 \paragraph{}Next, we attempted to replicate these experiments with the more complicated Human Driver Model.
@@
 \paragraph{}The models used within this paper necessarily make many simplifying assumptions, many of which vary somewhat between models.
-Because of this, it was often difficult to maintain stability of the systems in blended models, as one model may not fully satisfy the assumptions of the other under all parameters.
+Because of this, it was often difficult to maintain stability in the systems composed of a mixture of the two models, as one model may not fully satisfy the assumptions of the other under all parameters.
 However, within these restrictions, we were still able to construct a series of simulations with stable parameter values and model the reactions of the systems in reasonable ways.
